\chapter{Directives}
\index{directives}
\label{chap:Directives}
This chapter describes the syntax and behavior of OpenMP directives.

\begin{ccppspecific}
In C/C++, OpenMP directives are specified by using the \pcode{\#pragma} mechanism 
provided by the C and C++ standards.
\end{ccppspecific}

\begin{fortranspecific}
In Fortran, OpenMP directives are specified by using special comments that are
identified by unique sentinels. Also, a special comment form is available for 
conditional compilation.
\end{fortranspecific}

Compilers can therefore ignore OpenMP directives and conditionally compiled code if
support of the OpenMP API is not provided or enabled. A compliant implementation
must provide an option or interface that ensures that underlying support of all OpenMP
directives and OpenMP conditional compilation mechanisms is enabled. In the
remainder of this document, the phrase \emph{OpenMP compilation} is used to mean a
compilation with these OpenMP features enabled.

\begin{samepage}
\begin{fortranspecific}
\restrictions
The following restriction applies to all OpenMP directives:
\begin{itemize}
  \item OpenMP directives, except \code{simd} and any declarative directive,
 may not appear in pure procedures.
\item OpenMP directives may not appear in the WHERE and FORALL constructs.
\end{itemize}
\end{fortranspecific}
\end{samepage}


\input{directives/directives_common}
\input{directives/directives_parallel}
\input{directives/directives_teams}
\input{directives/directives_worksharing}
\input{directives/directives_loops}
\input{directives/directives_tasking}
\input{directives/directives_mm}
\input{directives/directives_device}
\input{directives/directives_combined}
\input{directives/directives_if}
\input{directives/directives_synchronization}
\input{directives/directives_cancellation}
\input{directives/directives_data_environment}
\input{directives/directives_nesting}

% This is the end of ch2-directives.tex of the OpenMP specification.

