% This is runtime_library.tex (Chapter 3) of the OpenMP specification.
% This is an included file. See the master file for more information.
%
% When editing this file:
%
%    1. To change formatting, appearance, or style, please edit openmp.sty.
%
%    2. Custom commands and macros are defined in openmp.sty.
%
%    3. Be kind to other editors -- keep a consistent style by copying-and-pasting to
%       create new content.
%
%    4. We use semantic markup, e.g. (see openmp.sty for a full list):
%         \code{}     % for bold monospace keywords, code, operators, etc.
%         \plc{}      % for italic placeholder names, grammar, etc.
%
%    5. There are environments that provide special formatting, e.g. language bars.
%       Please use them whereever appropriate.  Examples are:
%
%         \begin{fortranspecific}
%         This is text that appears enclosed in blue language bars for Fortran.
%         \end{fortranspecific}
%
%         \begin{note}
%         This is a note.  The "Note -- " header appears automatically.
%         \end{note}
%
%    6. Other recommendations:
%         Use the convenience macros defined in openmp.sty for the minor headers
%         such as Comments, Syntax, etc.
%
%         To keep items together on the same page, prefer the use of
%         \begin{samepage}.... Avoid \parbox for text blocks as it interrupts line numbering.
%         When possible, avoid \filbreak, \pagebreak, \newpage, \clearpage unless that's
%         what you mean. Use \needspace{} cautiously for troublesome paragraphs.
%
%         Avoid absolute lengths and measures in this file; use relative units when possible.
%         Vertical space can be relative to \baselineskip or ex units. Horizontal space
%         can be relative to \linewidth or em units.
%
%         Prefer \emph{} to italicize terminology, e.g.:
%             This is a \emph{definition}, not a placeholder.
%             This is a \plc{var-name}.
%

\chapter{Runtime Library Routines}
\index{runtime library routines}
\label{chap:Runtime Library Routines}
This chapter describes the OpenMP API runtime library routines and queryable runtime states, and is divided into the
following sections:

\begin{itemize}
\item Runtime library definitions
(\specref{sec:runtime library definitions}).

\item Execution environment routines that can be used to control and to query the parallel
execution environment
(\specref{sec:Execution Environment Routines}).

\item Lock routines that can be used to synchronize access to data
(\specref{sec:Lock Routines}).

\item Portable timer routines
(\specref{sec:Timing Routines}).

\item Device memory routines that can be used to allocate memory and
to manage pointers on target devices (\specref{sec:Device Memory Routines}).

%\item Runtime states (\specref{sec:runtimeStates}).

% \item OMPT Query Functions (\specref{sec:omptFunctions}).

\item Execution routines to control the application monitoring
(\specref{sec:control_tool})
\end{itemize}

Throughout this chapter, \plc{true} and \plc{false} are used as generic terms to simplify the
description of the routines.

\begin{samepage}
\begin{ccppspecific}
\plc{true} means a nonzero integer value and \plc{false} means an integer value of zero.
\end{ccppspecific}
\end{samepage}
\bigskip

\begin{samepage}
\begin{fortranspecific}
\plc{true} means a logical value of \code{.TRUE.} and \plc{false} means a logical value of \code{.FALSE.}.
\end{fortranspecific}
\end{samepage}
\bigskip

\begin{samepage}
\vspace{1\baselineskip}
\begin{fortranspecific}
\vspace{-1\baselineskip}
\restrictions

The following restriction applies to all OpenMP runtime library routines:

\begin{itemize}
\item OpenMP runtime library routines may not be called from \code{PURE} or \code{ELEMENTAL}
procedures.
\end{itemize}
\end{fortranspecific}
\end{samepage}











\section{Runtime Library Definitions}
\index{runtime library definitions}
\index{header files}
\index{include files}
\label{sec:runtime library definitions}
For each base language, a compliant implementation must supply a set of definitions for
the OpenMP API runtime library routines and the special data types of their parameters.
The set of definitions must contain a declaration for each OpenMP API runtime library
routine and a declaration for the \emph{simple lock}, \emph{nestable lock}, \emph{schedule}, and \emph{thread affinity
policy} data types. In addition, each set of definitions may specify other implementation
specific values.

\begin{ccppspecific}
The library routines are external functions with ``C'' linkage.

Prototypes for the C/C++ runtime library routines described in this chapter shall be
provided in a header file named \code{omp.h}. This file defines the following:

\begin{itemize}
\item The prototypes of all the routines in the chapter.

\item The type \code{omp_lock_t}.

\item The type \code{omp_nest_lock_t}.

\item The type \code{omp_lock_hint_t}.

\item The type \code{omp_sched_t}.

\item The type \code{omp_proc_bind_t}.

\item The type \code{omp_control_tool_t}.

\item The type \code{omp_control_tool_result_t}.

\item The type \code{omp_memspace_t}.

\item The type \code{omp_allocator_t}.

\item A global variable of type \code{const omp_memspace_t *} for each predefined memory space in \tabref{tab:Predefined Memory Spaces}.

\item A global variable of type \code{const omp_allocator_t *} for each predefined memory allocator in \tabref{tab:Predefined Allocators}.

A program that declares a new variable with the same identifier as one of the predefined allocators listed in \tabref{tab:Predefined Allocators} results in unspecified behavior.

\end{itemize}

%\newpage %% HACK
% was: C.1 on page 288
See \specref{sec:Example of the omp.h Header File} for an example of this file.

\end{ccppspecific}

\begin{fortranspecific}
The OpenMP Fortran API runtime library routines are external procedures. The return
values of these routines are of default kind, unless otherwise specified.

Interface declarations for the OpenMP Fortran runtime library routines described in this
chapter shall be provided in the form of a Fortran \code{include} file named \code{omp_lib.h} or
a Fortran~90 \code{module} named \code{omp_lib}. It is implementation defined whether the
\code{include} file or the \code{module} file (or both) is provided.

These files define the following:

\begin{itemize}
\item The interfaces of all of the routines in this chapter.

\item The \code{integer} \code{parameter} \code{omp_lock_kind}.

\item The \code{integer} \code{parameter} \code{omp_nest_lock_kind}.

\item The \code{integer} \code{parameter} \code{omp_lock_hint_kind}.

\item The \code{integer} \code{parameter} \code{omp_sched_kind}.

\item The \code{integer} \code{parameter} \code{omp_proc_bind_kind}.

\item The \code{integer} \code{parameter} \code{omp_memspace_kind}.

\item The \code{integer} \code{parameter} \code{omp_allocator_kind}.

\item The \code{integer} \code{parameter} \code{omp_alloctrait_key_kind}.

\item The \code{integer} \code{parameter} \code{omp_alloctrait_val_kind}.

\item An \code{integer} \code{parameter} variable of kind \code{omp_memspace_kind} for each predefined memory space in \tabref{tab:Predefined Memory Spaces}.

\item An \code{integer} \code{parameter} variable of kind \code{omp_allocator_kind} for each predefined memory allocator in \tabref{tab:Predefined Allocators}.

\item The \code{integer} \code{parameter} \code{openmp_version} with a value \plc{yyyymm} where \plc{yyyy}
and \plc{mm} are the year and month designations of the version of the OpenMP Fortran
API that the implementation supports. This value matches that of the C preprocessor
macro \code{_OPENMP}, when a macro preprocessor is supported (see
\specref{sec:Conditional Compilation}).
\end{itemize}

See \specref{sec:Example of an Interface Declaration include File}
and \specref{sec:Example of a Fortran Interface Declaration module}
for examples of these files.

It is implementation defined whether any of the OpenMP runtime library routines that
take an argument are extended with a generic interface so arguments of different \code{KIND}
type can be accommodated. See Appendix~\ref{sec:Example of a Generic Interface for a Library Routine}
for an example of such an extension.
\end{fortranspecific}

\input{runtime_library/runtime_library_execution}
\input{runtime_library/runtime_library_others}
\input{runtime_library/runtime_library_device_mem}
\input{runtime_library/runtime_library_mm}
\input{runtime_library/runtime_library_tools}

