% This is an included file. See the master file for more information.
%
% When editing this file:
%
%    1. To change formatting, appearance, or style, please edit openmp.sty.
%
%    2. Custom commands and macros are defined in openmp.sty.
%
%    3. Be kind to other editors -- keep a consistent style by copying-and-pasting to
%       create new content.
%
%    4. We use semantic markup, e.g. (see openmp.sty for a full list):
%         \code{}     % for bold monospace keywords, code, operators, etc.
%         \plc{}      % for italic placeholder names, grammar, etc.
%
%    5. There are environments that provide special formatting, e.g. language bars.
%       Please use them whereever appropriate.  Examples are:
%
%         \begin{fortranspecific}
%         This is text that appears enclosed in blue language bars for Fortran.
%         \end{fortranspecific}
%
%         \begin{note}
%         This is a note.  The "Note -- " header appears automatically.
%         \end{note}
%
%    6. Other recommendations:
%         Use the convenience macros defined in openmp.sty for the minor headers
%         such as Comments, Syntax, etc.
%
%         To keep items together on the same page, prefer the use of
%         \begin{samepage}.... Avoid \parbox for text blocks as it interrupts line numbering.
%         When possible, avoid \filbreak, \pagebreak, \newpage, \clearpage unless that's
%         what you mean. Use \needspace{} cautiously for troublesome paragraphs.
%
%         Avoid absolute lengths and measures in this file; use relative units when possible.
%         Vertical space can be relative to \baselineskip or ex units. Horizontal space
%         can be relative to \linewidth or em units.
%
%         Prefer \emph{} to italicize terminology, e.g.:
%             This is a \emph{definition}, not a placeholder.
%             This is a \plc{var-name}.
%


\section{Data Environment}
\label{sec:Data Environment}
\index{data environment}
This section presents directives and clauses for controlling data environments.

\subsection{Data-sharing Attribute Rules}
\label{subsec:Data-sharing Attribute Rules}
\index{data-sharing attribute rules}
\index{attributes, data-sharing}
This section describes how the data-sharing attributes of variables referenced in
data environments are determined. The following two cases are described separately:

\begin{itemize}
\item \specref{subsubsec:Variables Referenced in a Construct} describes the 
      data-sharing attribute rules for variables referenced in a construct.
\item \specref{subsubsec:Variables Referenced in a Region but not in a Construct} 
      describes the data-sharing attribute rules for variables referenced in a 
      region, but outside any construct.
\end{itemize}



\subsubsection{Variables Referenced in a Construct}
\label{subsubsec:Variables Referenced in a Construct}
The data-sharing attributes of variables that are referenced in a construct can be
\emph{predetermined}, \emph{explicitly determined}, or \emph{implicitly determined}, 
according to the rules outlined in this section.

Specifying a variable on a data-sharing attribute clause, except for the 
\code{private} clause, or \code{copyprivate} clause of an enclosed construct 
causes an implicit reference to the variable in the enclosing construct. 
Specifying a variable on a \code{map} clause of an enclosed construct may 
cause an implicit reference to the variable in the enclosing construct.
Such implicit references are also subject to the data-sharing attribute 
rules outlined in this section.

Certain variables and objects have \emph{predetermined} data-sharing 
attributes as follows:

\begin{ccppspecific}
\begin{itemize}
\item Variables that appear in \code{threadprivate} directives are threadprivate.
\item Variables with automatic storage duration that are declared in a scope 
      inside the construct are private.
\item Objects with dynamic storage duration are shared.
\item Static data members are shared.
\item The loop iteration variable(s) in the associated \plc{for-loop(s)} of a
      \code{for}, \code{parallel}~\code{for}, \code{taskloop}, or 
      \code{distribute} construct is (are) private.
\item The loop iteration variable in the associated \plc{for-loop} of a
      \code{simd} or \code{loop} construct with just one associated 
      \plc{for-loop} is linear with a \plc{linear-step} that is the increment 
      of the associated \plc{for-loop}.
\item The loop iteration variables in the associated \plc{for-loops} of a
      \code{simd} or \code{loop} construct with multiple associated 
      \plc{for-loops} are lastprivate.
\item Variables with static storage duration that are declared in a scope 
      inside the construct are shared.
\item If a list item in a \code{map} clause on the \code{target} construct 
      has a base pointer, and the base pointer is a scalar variable that 
      does not appear in a \code{map} clause on the construct, the base 
      pointer is firstprivate.
\end{itemize}
\end{ccppspecific}

\begin{fortranspecific}
\begin{itemize}
\item Variables and common blocks that appear in \code{threadprivate} directives
      are threadprivate.
\item The loop iteration variable(s) in the associated \plc{do-loop(s)} of a
      \code{do}, \code{parallel}~\code{do}, \code{taskloop}, or \code{distribute} 
      construct is (are) private.
\item The loop iteration variable in the associated \plc{do-loop} of a
      \code{simd} or \code{loop}  construct with just one associated 
      \plc{do-loop} is linear with a \plc{linear-step} that is the 
      increment of the associated \plc{do-loop}.
\item The loop iteration variables in the associated \plc{do-loops} of a
      \code{simd} or \code{loop} construct with multiple associated 
      \plc{do-loops} are lastprivate.
\item A loop iteration variable for a sequential loop in a \code{parallel} 
      or task generating construct is private in the innermost such construct 
      that encloses the loop.
\item Implied-do indices and \code{forall} indices are private.
\item Cray pointees have the same data-sharing attribute as the storage 
      with which their Cray pointers are associated.
\item Assumed-size arrays are shared.
\item An associate name preserves the association with the selector established 
      at the \code{ASSOCIATE} or \code{SELECT}~\code{TYPE} statement.
\end{itemize}
\end{fortranspecific}

Variables with predetermined data-sharing attributes may not be listed in 
data-sharing attribute clauses, except for the cases listed below. For these 
exceptions only, listing a predetermined variable in a data-sharing attribute 
clause is allowed and overrides the variable's predetermined data-sharing attributes.

\begin{ccppspecific}
\begin{itemize}
\item The loop iteration variable(s) in the associated \plc{for-loop(s)} of a 
      \code{for}, \code{parallel}~\code{for}, \code{taskloop}, or \code{distribute} 
      construct may be listed in a \code{private} or \code{lastprivate} clause.
\item The loop iteration variable in the associated \plc{for-loop} of a 
      \code{simd} construct with just one associated \plc{for-loop} may be 
      listed in a \code{private}, \code{lastprivate}, or \code{linear} clause 
      with a \plc{linear-step} that is the increment of the associated \plc{for-loop}.
\item The loop iteration variables in the associated \plc{for-loops} of a 
      \code{simd} construct with multiple associated \plc{for-loops} may be 
      listed in a \code{private} or \code{lastprivate} clause.
\item The loop iteration variable(s) in the associated \plc{for-loop(s)} of 
      a \code{loop} construct may be listed in a \code{private} clause.
\item Variables with \code{const}-qualified type with no mutable members may 
      be listed in a \code{firstprivate} clause, even if they are static data members.
\end{itemize}
\end{ccppspecific}

\begin{fortranspecific}
\begin{itemize}
\item The loop iteration variable(s) in the associated \plc{do-loop(s)} of a 
      \code{do}, \code{parallel}~\code{do}, \code{taskloop}, or \code{distribute}
      construct may be listed in a \code{private} or \code{lastprivate} clause.
\item The loop iteration variable in the associated \plc{d}o-loop of a \code{simd} 
      construct with just one associated \plc{do-loop} may be listed in a 
      \code{private}, \code{lastprivate}, or \code{linear} clause with a 
      \plc{linear-step} that is the increment of the associated loop.
\item The loop iteration variables in the associated \plc{do-loops} of a 
      \code{simd} construct with multiple associated \plc{do-loops} may be 
      listed in a \code{private} or \code{lastprivate} clause.
\item The loop iteration variable(s) in the associated \plc{do-loop(s)} of 
      a \code{loop} construct may be listed in a \code{private} clause.
\item Variables used as loop iteration variables in sequential loops in a 
      \code{parallel} or task generating construct may be listed in 
      data-sharing attribute clauses on the construct itself, and on
      enclosed constructs, subject to other restrictions.
\item Assumed-size arrays may be listed in a \code{shared} clause.
\end{itemize}
\end{fortranspecific}

Additional restrictions on the variables that may appear in individual clauses are
described with each clause in \specref{subsec:Data-Sharing Attribute Clauses}.

Variables with \emph{explicitly determined} data-sharing attributes are those that 
are referenced in a given construct and are listed in a data-sharing attribute clause 
on the construct.

Variables with \emph{implicitly determined} data-sharing attributes are those that 
are referenced in a given construct, do not have predetermined data-sharing 
attributes, and are not listed in a data-sharing attribute clause on the construct.

Rules for variables with \emph{implicitly determined} data-sharing attributes 
are as follows:

\begin{itemize}
\item In a \code{parallel}, \code{teams}, or task generating construct, the 
      data-sharing attributes of these variables are determined by the 
      \code{default} clause, if present (see \specref{subsubsec:default clause}).
\item In a \code{parallel} construct, if no \code{default} clause is present, 
      these variables are shared.
\item For constructs other than task generating constructs, if no \code{default} 
      clause is present, these variables reference the variables with the same 
      names that exist in the enclosing context.
\item In a \code{target} construct, variables that are not mapped after applying 
      data-mapping attribute rules (see 
      \specref{subsec:Data-Mapping Attribute Rules, Clauses, and Directives}) 
      are firstprivate.
\end{itemize}

\begin{cppspecific}
\begin{itemize}
\item In an orphaned task generating construct, if no \code{default} clause is 
      present, formal arguments passed by reference are firstprivate.
\end{itemize}
\end{cppspecific}

\begin{fortranspecific}
\begin{itemize}
\item In an orphaned task generating construct, if no \code{default} clause is 
      present, dummy arguments are firstprivate.
\end{itemize}
\end{fortranspecific}

\begin{itemize}
\item In a task generating construct, if no \code{default} clause is present, a 
      variable for which the data-sharing attribute is not determined by the rules 
      above and that in the enclosing context is determined to be shared by all 
      implicit tasks bound to the current team is shared.
\item In a task generating construct, if no \code{default} clause is present, a 
      variable for which the data-sharing attribute is not determined by the rules 
      above is firstprivate.
\end{itemize}

Additional restrictions on the variables for which data-sharing attributes cannot be
implicitly determined in a task generating construct are described in
\specref{subsubsec:firstprivate clause}.



\subsubsection{Variables Referenced in a Region but not in a Construct}
\label{subsubsec:Variables Referenced in a Region but not in a Construct}
The data-sharing attributes of variables that are referenced in a region, but not 
in a construct, are determined as follows:

\begin{ccppspecific}
\begin{itemize}
\item Variables with static storage duration that are declared in called routines 
      in the region are shared.
\item File-scope or namespace-scope variables referenced in called routines in the 
      region are shared unless they appear in a \code{threadprivate} directive.
\item Objects with dynamic storage duration are shared.
\item Static data members are shared unless they appear in a \code{threadprivate} 
      directive.
\item In C++, formal arguments of called routines in the region that are passed
      by reference have the same data-sharing attributes as the associated actual 
      arguments.
\item Other variables declared in called routines in the region are private.
\end{itemize}
\end{ccppspecific}

\begin{fortranspecific}
\begin{itemize}
\item Local variables declared in called routines in the region and that have the 
      \code{save} attribute, or that are data initialized, are shared unless they 
      appear in a \code{threadprivate} directive.
\item Variables belonging to common blocks, or accessed by host or use association, 
      and referenced in called routines in the region are shared unless they appear 
      in a \code{threadprivate} directive.
\item Dummy arguments of called routines in the region that have the
      \code{VALUE} attribute are private.
\item Dummy arguments of called routines in the region that do not have the
      \code{VALUE} attribute are private if the associated actual argument is not 
      shared.
\item Dummy arguments of called routines in the region that do not have the
      \code{VALUE} attribute are shared if the actual argument is shared and 
      it is a scalar variable, structure, an array that is not a pointer or
      assumed-shape array, or a simply contiguous array section.  Otherwise, the
      data-sharing attribute of the dummy argument is implementation-defined if
      the associated actual argument is shared.
\item Cray pointees have the same data-sharing attribute as the storage with 
      which their Cray pointers are associated.
\item Implied-do indices, \code{forall} indices, and other local variables 
      declared in called routines in the region are private.
\end{itemize}
\end{fortranspecific}



\subsection{\hcode{threadprivate} Directive}
\index{threadprivate@{\code{threadprivate}}}
\index{directives!threadprivate@{\code{threadprivate}}}
\label{subsec:threadprivate Directive}
\summary
The \code{threadprivate} directive specifies that variables are replicated, with 
each thread having its own copy. The \code{threadprivate} directive is a declarative 
directive.

\syntax
\begin{ccppspecific}
The syntax of the \code{threadprivate} directive is as follows:

\begin{ompcPragma}
#pragma omp threadprivate(\plc{list}) \plc{new-line}
\end{ompcPragma}

where \plc{list} is a comma-separated list of file-scope, namespace-scope, or static
block-scope variables that do not have incomplete types.
\end{ccppspecific}

\begin{fortranspecific}
The syntax of the \code{threadprivate} directive is as follows:

\begin{ompfPragma}
!$omp threadprivate(\plc{list})
\end{ompfPragma}

where \plc{list} is a comma-separated list of named variables and named common blocks.
Common block names must appear between slashes.
\end{fortranspecific}

\descr
Each copy of a threadprivate variable is initialized once, in the manner specified 
by the program, but at an unspecified point in the program prior to the first 
reference to that copy. The storage of all copies of a threadprivate variable is 
freed according to how static variables are handled in the base language, but at 
an unspecified point in the program.

A program in which a thread references another thread's copy of a threadprivate 
variable is non-conforming.

The content of a threadprivate variable can change across a task scheduling point 
if the executing thread switches to another task that modifies the variable. For 
more details on task scheduling, see \specref{sec:Execution Model} and
\specref{sec:Tasking Constructs}.

In \code{parallel} regions, references by the master thread will be to the copy of the
variable in the thread that encountered the \code{parallel} region.

During a sequential part references will be to the initial thread's copy of the 
variable. The values of data in the initial thread's copy of a threadprivate 
variable are guaranteed to persist between any two consecutive references to 
the variable in the program.

The values of data in the threadprivate variables of non-initial threads
are guaranteed to persist between two consecutive active \code{parallel}
regions only if all of the following conditions hold:

\begin{itemize} 
\item Neither \code{parallel} region is nested inside another explicit 
      \code{parallel} region;
\item The number of threads used to execute both \code{parallel} regions is the same;
\item The thread affinity policies used to execute both \code{parallel} 
      regions are the same;
\item The value of the \plc{dyn-var} internal control variable in the enclosing 
      task region is \plc{false} at entry to both \code{parallel} regions; and
\item Neither the \code{omp_pause_resource} nor \code{omp_pause_resource_all} 
      routine is called.
\end{itemize}

If these conditions all hold, and if a threadprivate variable is referenced in 
both regions, then threads with the same thread number in their respective 
regions will reference the same copy of that variable.

\begin{ccppspecific}
If the above conditions hold, the storage duration, lifetime, and value of a 
thread's copy of a threadprivate variable that does not appear in any 
\code{copyin} clause on the second region will be retained. Otherwise, the 
storage duration, lifetime, and value of a thread's copy of the variable in 
the second region is unspecified.
\end{ccppspecific}

\begin{fortranspecific}
If the above conditions hold, the definition, association, or allocation 
status of a thread's copy of a threadprivate variable or a variable in a 
threadprivate common block that is not affected by any \code{copyin} clause 
that appears on the second region (a variable is affected by a \code{copyin} 
clause if the variable appears in the \code{copyin} clause or it is in a 
common block that appears in the \code{copyin} clause) will be retained.
Otherwise, the definition and association status of a thread's copy of the 
variable in the second region are undefined, and the allocation status of 
an allocatable variable will be implementation defined.

If a threadprivate variable or a variable in a threadprivate common block is
not affected by any \code{copyin} clause that appears on the first \code{parallel} 
region in which it is referenced, the thread's copy of the variable inherits the
declared type parameter and the default parameter values from the original variable.  
The variable or any subobject of the variable is initially defined or
undefined according to the following rules:

\begin{itemize}
\item If it has the \code{ALLOCATABLE} attribute, each copy created will have 
      an initial allocation status of unallocated;
\item If it has the \code{POINTER} attribute:

\begin{itemize}
\item If it has an initial association status of disassociated, either through 
      explicit initialization or default initialization, each copy created will 
      have an association status of disassociated;
\item Otherwise, each copy created will have an association status of undefined.
\end{itemize}

\item If it does not have either the \code{POINTER} or the \code{ALLOCATABLE} 
      attribute:

\begin{samepage}
\begin{itemize}
\item If it is initially defined, either through explicit initialization or default
      initialization, each copy created is so defined;
\item Otherwise, each copy created is undefined.
\end{itemize}
\end{samepage}
\end{itemize}
\end{fortranspecific}

\begin{ccppspecific}
The address of a threadprivate variable is not an address constant.
\end{ccppspecific}

\begin{cppspecific}
The order in which any constructors for different threadprivate variables 
of class type are called is unspecified. The order in which any destructors 
for different threadprivate variables of class type are called is unspecified.
\end{cppspecific}

\restrictions
The restrictions to the \code{threadprivate} directive are as follows:

\begin{itemize}
\item A threadprivate variable must not appear in any clause except the 
      \code{copyin}, \code{copyprivate}, \code{schedule}, \code{num_threads}, 
      \code{thread_limit}, and \code{if} clauses.
\item A program in which an untied task accesses threadprivate storage is 
      non-conforming.

\begin{ccppspecific}
\item If the value of a variable referenced in an explicit initializer of a 
      threadprivate variable is modified prior to the first reference to any 
      instance of the threadprivate variable, then the behavior is unspecified.
\item A variable that is part of another variable (as an array or structure 
      element) cannot appear in a \code{threadprivate} clause unless it is a 
      static data member of a C++ class.
\item A \code{threadprivate} directive for file-scope variables must appear 
      outside any definition or declaration, and must lexically precede all 
      references to any of the variables in its list.
\item A \code{threadprivate} directive for namespace-scope variables must 
      appear outside any definition or declaration other than the namespace 
      definition itself, and must lexically precede all references to any of 
      the variables in its list.
\item Each variable in the list of a \code{threadprivate} directive at file, 
      namespace, or class scope must refer to a variable declaration at file, 
      namespace, or class scope that lexically precedes the directive.
\item A \code{threadprivate} directive for static block-scope variables must 
      appear in the scope of the variable and not in a nested scope. The 
      directive must lexically precede all references to any of the variables 
      in its list.
\item Each variable in the list of a \code{threadprivate} directive in block 
      scope must refer to a variable declaration in the same scope that lexically 
      precedes the directive. The variable declaration must use the static 
      storage-class specifier.
\item If a variable is specified in a \code{threadprivate} directive in one 
      translation unit, it must be specified in a \code{threadprivate} directive 
      in every translation unit in which it is declared.
\end{ccppspecific}

\begin{cppspecific}
\item A \code{threadprivate} directive for static class member variables must 
      appear in the class definition, in the same scope in which the member 
      variables are declared, and must lexically precede all references to 
      any of the variables in its list.
\item A threadprivate variable must not have an incomplete type or a reference type.
\item A threadprivate variable with class type must have:

\begin{itemize}
\item An accessible, unambiguous default constructor in the case of default 
      initialization without a given initializer;
\item An accessible, unambiguous constructor that accepts the given argument
      in the case of direct initialization; and
\item An accessible, unambiguous copy constructor in the case of copy initialization 
      with an explicit initializer.
\end{itemize}
\end{cppspecific}
\end{itemize}

\begin{fortranspecific}
\begin{itemize}
\item A variable that is part of another variable (as an array, structure element 
      or type parameter inquiry) cannot appear in a \code{threadprivate} clause.
\item The \code{threadprivate} directive must appear in the declaration section 
      of a scoping unit in which the common block or variable is declared.
\item If a \code{threadprivate} directive that specifies a common block name appears 
      in one program unit, then such a directive must also appear in every other 
      program unit that contains a \code{COMMON} statement that specifies the same 
      name. It must appear after the last such \code{COMMON} statement in the 
      program unit.
\item If a threadprivate variable or a threadprivate common block is declared
      with the \code{BIND} attribute, the corresponding C entities must also be 
      specified in a \code{threadprivate} directive in the C program.
\item A blank common block cannot appear in a \code{threadprivate} directive.
\item A variable can only appear in a \code{threadprivate} directive in the 
      scope in which it is declared. It must not be an element of a common block 
      or appear in an \code{EQUIVALENCE} statement.
\item A variable that appears in a \code{threadprivate} directive must be declared 
      in the scope of a module or have the \code{SAVE} attribute, either explicitly 
      or implicitly.
\end{itemize}
\end{fortranspecific}

\crossreferences
\begin{itemize}
\item \plc{dyn-var} ICV, see
\specref{sec:Internal Control Variables}.

\item Number of threads used to execute a \code{parallel} region, see
\specref{subsec:Determining the Number of Threads for a parallel Region}.

\item \code{copyin} clause, see
\specref{subsubsec:copyin clause}.
\end{itemize}



\subsection{List Item Privatization}
\label{subsec:List Item Privatization}
\index{list item privatization}

For any construct, a list item that appears in a data-sharing attribute
clause, including a reduction clause, may be privatized. Each task that
references a privatized list item in any statement in the construct receives
at least one new list item if the construct has one or more associated loops,
and otherwise each such task receives one new list item.  Each SIMD lane used
in a \code{simd} construct that references a privatized list item in any
statement in the construct receives at least one new list item.
Language-specific attributes for new list items are derived from the
corresponding original list item. Inside the construct, all references to the
original list item are replaced by references to a new list item received by
the task or SIMD lane.  

If the construct has one or more associated loops, within the same logical
iteration of the loop(s) the same new list item replaces all references to the
original list item. For any two logical iterations, if the references to the
original list item are replaced by the same list item then the logical
iterations must execute in some sequential order.

In the rest of the region, it is unspecified whether references are
to a new list item or the original list item. Therefore, if an attempt is made
to reference the original item, its value after the region is also
unspecified. If a task or a SIMD lane does not reference a privatized list
item, it is unspecified whether the task or SIMD lane receives a new list
item.

The value and/or allocation status of the original list item will change only:

\begin{itemize}
\item If accessed and modified via pointer;
\item If possibly accessed in the region but outside of the construct;
\item As a side effect of directives or clauses; or

\begin{fortranspecific}
\item If accessed and modified via construct association.
\end{fortranspecific}
\end{itemize}

\begin{cppspecific}
If the construct is contained in a member function, it is unspecified
anywhere in the region if accesses through the implicit \code{this}
pointer refer to the new list item or the original list item.
\end{cppspecific}

\begin{ccppspecific}
A new list item of the same type, with automatic storage duration, is allocated 
for the construct. The storage and thus lifetime of these list items last until 
the block in which they are created exits. The size and alignment of the new list 
item are determined by the type of the variable. This allocation occurs once for 
each task generated by the construct and once for each SIMD lane used by the construct.

The new list item is initialized, or has an undefined initial value, as if it 
had been locally declared without an initializer.
\end{ccppspecific}

\begin{cppspecific}
If the type of a list item is a reference to a type \plc{T} then the type will 
be considered to be \plc{T} for all purposes of this clause.

The order in which any default constructors for different private variables of 
class type are called is unspecified. The order in which any destructors for 
different private variables of class type are called is unspecified.
\end{cppspecific}

\begin{fortranspecific}
If any statement of the construct references a list item, a new list
item of the same type and type parameters is allocated. This
allocation occurs once for each task generated by the construct and
once for each SIMD lane used by the construct. The initial value of
the new list item is undefined. The initial status of a private
pointer is undefined.

For a list item or the subobject of a list item with the \code{ALLOCATABLE} attribute:

\begin{itemize}
\item If the allocation status is unallocated, the new list item or the subobject
      of the new list item will have an initial allocation status of unallocated;
\item If the allocation status is allocated, the new list item or the subobject of
      the new list item will have an initial allocation status of allocated; and
\item If the new list item or the subobject of the new list item is an array, its 
      bounds will be the same as those of the original list item or the subobject 
      of the original list item.
\end{itemize}

A privatized list item may be storage-associated with other variables when the 
data-sharing attribute clause is encountered. Storage association may exist
because of constructs such as \code{EQUIVALENCE} or \code{COMMON}. If \plc{A}
is a variable that is privatized by a construct and \plc{B} is a variable that 
is storage-associated with \plc{A}, then:

\begin{itemize}
\item The contents, allocation, and association status of \plc{B} are undefined 
      on entry to the region;
\item Any definition of \plc{A}, or of its allocation or association status, 
      causes the contents, allocation, and association status of \plc{B} to 
      become undefined; and
\item Any definition of \plc{B}, or of its allocation or association status, 
      causes the contents, allocation, and association status of \plc{A} to 
      become undefined.
\end{itemize}

A privatized list item clause may be a selector of an \code{ASSOCIATE} or
\code{SELECT TYPE} construct. If the construct association is established
prior to a \code{parallel} region, the association between the associate name
and the original list item will be retained in the region.

Finalization of a list item of a finalizable type or subojects of a
list item of a finalizable type occurs at the end of the region. The
order in which any final subroutines for different variables of a
finalizable type are called is unspecified.
\end{fortranspecific}

If a list item appears in both \code{firstprivate} and \code{lastprivate} 
clauses, the update required for \code{lastprivate} occurs after all the 
initializations for \code{firstprivate}.

\restrictions
The following restrictions apply to any list item that is privatized unless
otherwise stated for a given data-sharing attribute clause:

\begin{itemize}
\begin{cspecific}
\item A variable that is part of another variable (as an array or structure
      element) cannot be privatized.
\end{cspecific}

\begin{cppspecific}
\item A variable that is part of another variable (as an array or structure 
      element) cannot be privatized except if the data-sharing attribute clause 
      is associated with a construct within a class non-static member function 
      and the variable is an accessible data member of the object for which the 
      non-static member function is invoked.
\item A variable of class type (or array thereof) that is privatized requires
      an accessible, unambiguous default constructor for the class type.
\end{cppspecific}

\begin{ccppspecific}
\item A variable that is privatized must not have a \code{const}-qualified type
      unless it is of class type with a \code{mutable} member. This restriction 
      does not apply to the \code{firstprivate} clause.
\item A variable that is privatized must not have an incomplete type or be a 
      reference to an incomplete type. 
\end{ccppspecific}

\begin{fortranspecific}
\item A variable that is part of another variable (as an array or structure 
      element) cannot be privatized.
\item A variable that is privatized must either be definable, or an allocatable 
      variable. This restriction does not apply to the \code{firstprivate} clause.
\item Variables that appear in namelist statements, in variable format expressions, 
      and in expressions for statement function definitions, may not be privatized.
\item Pointers with the \code{INTENT(IN)} attribute may not appear be privatized. 
      This restriction does not apply to the \code{firstprivate} clause.
\item Assumed-size arrays may not be privatized in a \code{target},
      \code{teams}, or \code{distribute} construct.
\end{fortranspecific}
\end{itemize}



\subsection{Data-Sharing Attribute Clauses}
\label{subsec:Data-Sharing Attribute Clauses}
\index{data-sharing attribute clauses}
\index{attribute clauses}
\index{clauses!data-sharing}
\index{clauses!attribute data-sharing}
Several constructs accept clauses that allow a user to control the data-sharing
attributes of variables referenced in the construct.  Not all of the clauses
listed in this section are valid on all directives. The set of clauses that is
valid on a particular directive is described with the directive.

Most of the clauses accept a comma-separated list of list items (see
\specref{sec:Directive Format}). All list items that appear in a clause must be 
visible, according to the scoping rules of the base language. With the exception 
of the \code{default} clause, clauses may be repeated as needed. A list item may 
not appear in more than one clause on the same directive, except that it may be 
specified in both \code{firstprivate} and \code{lastprivate} clauses.

The reduction data-sharing attribute clauses are explained in 
\specref{subsec:Reduction Clauses and Directives}.

\begin{cppspecific}
If a variable referenced in a data-sharing attribute clause has a type derived from a
template, and the program does not other wise reference that variable then any
behavior related to that variable is unspecified.
\end{cppspecific}

\begin{fortranspecific}
When a named common block appears in a \code{private}, \code{firstprivate},
\code{lastprivate}, or \code{shared} clause of a directive, none of its members 
may be declared in another data-sharing attribute clause in that directive. When 
individual members of a common block appear in a \code{private}, \code{firstprivate},
\code{lastprivate}, \code{reduction}, or \code{linear} clause of a directive, the 
storage of the specified variables is no longer Fortran associated with the storage 
of the common block itself. 
\end{fortranspecific}



\subsubsection{\hcode{default} Clause}
\label{subsubsec:default clause}
\index{default@{\code{default}}}
\index{clauses!default@{\code{default}}}
\summary
The \code{default} clause explicitly determines the data-sharing attributes 
of variables that are referenced in a \code{parallel}, \code{teams}, or task 
generating construct and would otherwise be implicitly determined (see
\specref{subsubsec:Variables Referenced in a Construct}).

\syntax
\begin{ccppspecific}
The syntax of the \code{default} clause is as follows:

\begin{ompSyntax}
default(shared \textnormal{|} none)
\end{ompSyntax}
\end{ccppspecific}

\begin{fortranspecific}
The syntax of the \code{default} clause is as follows:

\begin{ompSyntax}
default(private \textnormal{|} firstprivate \textnormal{|} shared \textnormal{|} none)
\end{ompSyntax}
\end{fortranspecific}

\descr
The \code{default(shared)} clause causes all variables referenced in the construct 
that have implicitly determined data-sharing attributes to be shared.

\begin{fortranspecific}
The \code{default(firstprivate)} clause causes all variables in the construct that 
have implicitly determined data-sharing attributes to be firstprivate.

The \code{default(private)} clause causes all variables referenced in the construct 
that have implicitly determined data-sharing attributes to be private.
\end{fortranspecific}

The \code{default(none)} clause requires that each variable that is referenced 
in the construct, and that does not have a predetermined data-sharing attribute, 
must have its data-sharing attribute explicitly determined by being listed in a 
data-sharing attribute clause.

\restrictions
The restrictions to the \code{default} clause are as follows:

\begin{itemize}
\item Only a single \code{default} clause may be specified on a
      \code{parallel}, \code{task}, \code{taskloop} or \code{teams} directive.
\end{itemize}



\subsubsection{\hcode{shared} Clause}
\label{subsubsec:shared clause}
\index{shared@{\code{shared}}}
\index{clauses!shared@{\code{shared}}}
\summary
The \code{shared} clause declares one or more list items to be shared by 
tasks generated by a \code{parallel}, \code{teams}, or task generating construct.

\syntax
The syntax of the \code{shared} clause is as follows:

\begin{ompSyntax}
shared(\plc{list})
\end{ompSyntax}

\descr
All references to a list item within a task refer to the storage area of the 
original variable at the point the directive was encountered.

The programmer must ensure, by adding proper synchronization, that storage 
shared by an explicit task region does not reach the end of its lifetime before
the explicit task region completes its execution.


\begin{fortranspecific}
The association status of a shared pointer becomes undefined upon entry to and 
on exit from the \code{parallel}, \code{teams}, or task generating construct if 
it is associated with a target or a  subobject of a target that is in a \code{private},
\code{firstprivate}, \code{lastprivate}, \code{reduction} clause, 
\code{task_redution} or \code{in_reduction} in the construct.


\begin{note}
Passing a shared variable to a procedure may result in the use of temporary 
storage in place of the actual argument when the corresponding dummy argument 
does not have the \code{VALUE} or \code{CONTIGUOUS} attribute and its 
data-sharing attribute is implementation-defined as per the rules in
\specref{subsubsec:Variables Referenced in a Region but not in a Construct}.
These conditions effectively result in references to, and definitions of, the
temporary storage during the procedure reference.  Furthermore, the value of
the shared variable is copied into the intervening temporary storage before 
the procedure reference when the dummy argument does not have the \code{INTENT(OUT)}
attribute, and is copied out of the temporary storage into the shared variable when
the dummy argument does not have the \code{INTENT(IN)} attribute.  Any
references to (or definitions of) the shared storage that is associated with
the dummy argument by any other task must be synchronized with
the procedure reference to avoid possible race conditions.
\end{note}
\end{fortranspecific}

\restrictions
The restrictions for the \code{shared} clause are as follows:

\begin{itemize}
\begin{cspecific}
\item A variable that is part of another variable (as an array or structure 
      element) cannot appear in a \code{shared} clause.
\end{cspecific}

\begin{cppspecific}
\item A variable that is part of another variable (as an array or structure
      element) cannot appear in a \code{shared} clause except if the \code{shared}
      clause is associated with a construct within a class non-static member
      function and the variable is an accessible data member of the object for
      which the non-static member function is invoked.
\end{cppspecific}

\begin{fortranspecific}
\item A variable that is part of another variable (as an array, structure element 
      or type parameter inquiry) cannot appear in a \code{shared} clause.
\end{fortranspecific}

\end{itemize}



\subsubsection{\hcode{private} Clause}
\index{private@{\code{private}}}
\index{clauses!private@{\code{private}}}
\label{subsubsec:private clause}
\summary
The \code{private} clause declares one or more list items to be private to a task 
or to a SIMD lane.

\syntax
The syntax of the \code{private} clause is as follows:

\begin{ompSyntax}
private(\plc{list})
\end{ompSyntax}

\descr

The \code{private} clause specifies that its list items are to be
privatized according to \specref{subsec:List Item Privatization}. Each task
or SIMD lane that references a list item in the construct receives only one
new list item, unless the construct has one or more associated loops and the
\code{order(concurrent)} clause is also present.

List items that appear in a \code{private}, \code{firstprivate}, or
\code{reduction} clause in a \code{parallel} construct may also appear
in a \code{private} clause in an enclosed \code{parallel}, worksharing, 
\code{loop}, \code{task}, \code{taskloop}, \code{simd}, or \code{target} construct.

List items that appear in a \code{private} or \code{firstprivate} clause 
in a \code{task} or \code{taskloop} construct may also appear in a \code{private}
clause in an enclosed \code{parallel}, \code{loop}, \code{task}, \code{taskloop}, 
\code{simd}, or \code{target} construct.

List items that appear in a \code{private}, \code{firstprivate},
\code{lastprivate}, or \code{reduction} clause in a worksharing
construct may also appear in a \code{private} clause in an enclosed
\code{parallel}, \code{loop}, \code{task}, \code{simd}, or \code{target} construct.

List items that appear in a \code{private} clause on a \code{loop} construct
may also appear in a \code{private} clause in an enclosed \code{loop}, 
\code{parallel}, or \code{simd} construct.

\restrictions
The restrictions to the \code{private} clause are as specified in
Section~\ref{subsec:List Item Privatization}.

\crossreferences
\begin{itemize}
\item List Item Privatization, see \specref{subsec:List Item Privatization}.
\end{itemize}



\subsubsection{\hcode{firstprivate} Clause}
\label{subsubsec:firstprivate clause}
\index{firstprivate@{\code{firstprivate}}}
\index{clauses!firstprivate@{\code{firstprivate}}}
\summary
The \code{firstprivate} clause declares one or more list items to be private to a task, and
initializes each of them with the value that the corresponding original item has when the
construct is encountered.

\syntax
The syntax of the \code{firstprivate} clause is as follows:

\begin{ompSyntax}
firstprivate(\plc{list})
\end{ompSyntax}

\descr
The \code{firstprivate} clause provides a superset of the functionality provided by the
\code{private} clause.

A list item that appears in a \code{firstprivate} clause is subject to the \code{private} clause
semantics described in
\specref{subsubsec:private clause},
except as noted. In addition, the
new list item is initialized from the original list item existing before the construct. The
initialization of the new list item is done once for each task that references the list item
in any statement in the construct. The initialization is done prior to the execution of the
construct.

For a \code{firstprivate} clause on a \code{parallel}, \code{task},
\code{taskloop}, \code{target}, or \code{teams} construct, the initial
value of the new list item is the value of the original list item that
exists immediately prior to the construct in the task region where the
construct is encountered unless otherwise specified. For a
\code{firstprivate} clause on a worksharing construct, the initial
value of the new list item for each implicit task of the threads that
execute the worksharing construct is the value of the original list
item that exists in the implicit task immediately prior to the point
in time that the worksharing construct is encountered unless otherwise
specified.

To avoid race conditions, concurrent updates of the original list item must be
synchronized with the read of the original list item that occurs as a result of the
\code{firstprivate} clause.

\begin{ccppspecific}
For variables of non-array type, the initialization occurs by copy assignment. For an
array of elements of non-array type, each element is initialized as if by assignment from
an element of the original array to the corresponding element of the new array.
\end{ccppspecific}
%
\begin{cppspecific}
For each variable of class type:
\begin{itemize}
\item If the \code{firstprivate} clause is not on a \code{target} construct then a copy constructor is invoked to perform the initialization;
\item If the \code{firstprivate} clause is on a \code{target} construct then it is unspecified how many copy constructors, if any, are invoked.
\end{itemize}
If copy constructors are called, the order in which copy constructors for different variables of class type are called is unspecified.
\end{cppspecific}
%
\begin{fortranspecific}
If the original list item does not have the \code{POINTER} attribute,
initialization of the new list items occurs as if by intrinsic
assignment unless the list item has a type bound procedure as a
defined assignment.  If the original list item that does not have the
\code{POINTER} attribute has the allocation status of unallocated, the
new list items will have the same status.

If the original list item has the \code{POINTER} attribute, the new list items receive the same
association status of the original list item as if by pointer assignment.
\end{fortranspecific}
%
\restrictions
The restrictions to the \code{firstprivate} clause are as follows:

\begin{itemize}
\item A list item that is private within a \code{parallel} region must
not appear in a \code{firstprivate} clause on a worksharing construct
if any of the worksharing regions arising
from the worksharing construct ever bind to any of the
\code{parallel} regions arising from the \code{parallel} construct.

\item A list item that is private within a \code{teams} region must not appear in a
\code{firstprivate} clause on a \code{distribute} construct if any of the \code{distribute}
regions arising from the \code{distribute} construct ever bind to any of the \code{teams}
regions arising from the \code{teams} construct.

\item A list item that appears in a \code{reduction} clause of a \code{parallel}
construct must not appear in a \code{firstprivate} clause on a worksharing, \code{task},
or \code{taskloop} construct if any of the worksharing or \code{task} regions arising from
the worksharing, \code{task}, or \code{taskloop} construct ever bind to any of the
\code{parallel} regions arising from the \code{parallel} construct.

\item A list item that appears in a \code{reduction} clause of a \code{teams} construct must not
appear in a \code{firstprivate} clause on a \code{distribute} construct if any of the
\code{distribute} regions arising from the \code{distribute} construct ever bind to any of
the \code{teams} regions arising from the \code{teams} construct.

\item A list item that appears in a \code{reduction} clause of a worksharing construct must not
appear in a \code{firstprivate} clause in a \code{task} construct encountered during execution
of any of the worksharing regions arising from the worksharing construct.

\begin{cppspecific}
\item A variable of class type (or array thereof) that appears in a \code{firstprivate} clause
requires an accessible, unambiguous copy constructor for the class type.
\end{cppspecific}
%
\begin{ccppspecific}
\item If a list item in a \code{firstprivate} clause on a worksharing
construct has a reference type then it must bind to the same object for all threads of the team.
\end{ccppspecific}
%
\begin{fortranspecific}
\item If the list item is a polymorphic variable with the \code{ALLOCATABLE} attribute, the behavior is unspecified.
\end{fortranspecific}
\end{itemize}










\subsubsection{\hcode{lastprivate} Clause}
\index{lastprivate@{\code{lastprivate}}}
\index{clauses!lastprivate@{\code{lastprivate}}}
\label{subsubsec:lastprivate clause}
\summary
The \code{lastprivate} clause declares one or more list items to be private to an implicit
task or to a SIMD lane, and causes the corresponding original list item to be updated
after the end of the region.

\syntax
The syntax of the \code{lastprivate} clause is as follows:

\begin{ompSyntax}
lastprivate(\plc{[ lastprivate-modifier}:\plc{] list})
\end{ompSyntax}

where \plc{lastprivate-modifier} is:
\begin{indentedcodelist}
conditional
\end{indentedcodelist}

\descr
The \code{lastprivate} clause provides a superset of the functionality provided by the
\code{private} clause.

A list item that appears in a \code{lastprivate} clause is subject to the \code{private} clause
semantics described in
\specref{subsubsec:private clause}.
In addition, when a
\code{lastprivate} clause without the \code{conditional} modifier appears on a directive,
the value of each new list item from the sequentially last iteration
of the associated loops, or the lexically last \code{section} construct, is assigned to the
original list item. When the \code{conditional} modifier appears on the clause,
if an assignment to a list item is encountered in the construct then the
original list item is assigned the value that is assigned to the new list item
in the sequentially last iteration or lexically last section in which such an
assignment is encountered.


\begin{ccppspecific}
For an array of elements of non-array type, each element is assigned to the
corresponding element of the original array.
\end{ccppspecific}
%\bigskip
%
\begin{fortranspecific}
If the original list item does not have the \code{POINTER} attribute,
its update occurs as if by intrinsic assignment unless it has a type
bound procedure as a defined assignment.

If the original list item has the \code{POINTER} attribute, its update occurs as if by pointer
assignment.
\end{fortranspecific}

When the \code{conditional} modifier does not appear on the \code{lastprivate} clause, list items that are not
assigned a value by the sequentially last iteration of the loops, or by the
lexically last \code{section} construct, have unspecified values after the
construct.  Unassigned subcomponents also have unspecified values after the
construct.

If the \code{lastprivate} clause is used on a construct to which neither
the \code{nowait} nor the \code{nogroup} clauses are applied, the original list item
becomes defined at the end of the construct.  To avoid race conditions,
concurrent reads or updates of the original list item must be synchronized with
the update of the original list item that occurs as a result of the
\code{lastprivate} clause.

Otherwise, If the \code{lastprivate} clause is used on a construct to which
the \code{nowait} or the \code{nogroup} clauses are applied, accesses to the original
list item may create a data race.  To avoid this, if an assignment to the
original list item occurs then synchronization must be inserted to ensure that
the assignment completes and the original list item is flushed to memory.

If a list item that appears in a \code{lastprivate} clause with the
\code{conditional} modifier is modified in the region by an assignment
outside the construct or not to the list item then the value assigned to
the original list item is unspecified.

\restrictions
The restrictions to the \code{lastprivate} clause are as follows:

\begin{itemize}
\item A list item that is private within a \code{parallel} region, or
that appears in the \code{reduction} clause of a \code{parallel}
construct, must not appear in a \code{lastprivate} clause on a
worksharing construct if any of the corresponding
worksharing regions ever binds to any of the corresponding
\code{parallel} regions.

\item A list item that appears in a \code{lastprivate} clause with the
\code{conditional} modifier must be a scalar variable.

\begin{cppspecific}
\item A variable of class type (or array thereof) that appears in a \code{lastprivate} clause
requires an accessible, unambiguous default constructor for the class type, unless the
list item is also specified in a \code{firstprivate} clause.

\item A variable of class type (or array thereof) that appears in a \code{lastprivate} clause
requires an accessible, unambiguous copy assignment operator for the class type. The
order in which copy assignment operators for different variables of class type are
called is unspecified.
\end{cppspecific}
%
\begin{ccppspecific}
\item If a list item in a \code{lastprivate} clause on a worksharing
construct has a reference type then it must bind to the same object for all threads of the team.
\end{ccppspecific}
%
\begin{fortranspecific}
\item A variable that appears in a \code{lastprivate} clause must be definable.

\item If the original list item has the \code{ALLOCATABLE} attribute, the
    corresponding list item whose value is assigned to the original list item must have an allocation status of allocated upon exit from
    the sequentially last iteration or lexically last \code{section} construct.

\item If the list item is a polymorphic variable with the \code{ALLOCATABLE} attribute, the behavior is unspecified.
\end{fortranspecific}
\end{itemize}










\subsubsection{\hcode{linear} Clause}
\index{linear@{\code{linear}}}
\index{clauses!linear@{\code{linear}}}
\label{subsubsec:linear clause}
\summary
The \code{linear} clause declares one or more list items to be private to a SIMD lane and to
have a linear relationship with respect to the iteration space of a loop.

\syntax


\begin{cspecific}
The syntax of the \code{linear} clause is as follows:
\begin{ompSyntax}
linear(\plc{linear-list[ }:\plc{ linear-step]})
\end{ompSyntax}
where \plc{linear-list} is one of the following
% 
\begin{indentedcodelist}
\plc{list}
\plc{modifier}(\plc{list})
\end{indentedcodelist}
where  \plc{modifier} is one of the following:
% 
\begin{indentedcodelist}
val
\end{indentedcodelist}
\end{cspecific}
%
\begin{cppspecific}
The syntax of the \code{linear} clause is as follows:
\begin{ompSyntax}
linear(\plc{linear-list[ }:\plc{ linear-step]})
\end{ompSyntax}
where \plc{linear-list} is one of the following
% 
\begin{indentedcodelist}
\plc{list}
\plc{modifier}(\plc{list})
\end{indentedcodelist}
where  \plc{modifier} is one of the following:
% 
\begin{indentedcodelist}
ref
val
uval
\end{indentedcodelist}
\end{cppspecific}

\begin{fortranspecific}
The syntax of the \code{linear} clause is as follows:
\begin{ompSyntax}
linear(\plc{linear-list[ }:\plc{ linear-step]})
\end{ompSyntax}
where \plc{linear-list} is one of the following
% 
\begin{indentedcodelist}
\plc{list}
\plc{modifier}(\plc{list})
\end{indentedcodelist}
where  \plc{modifier} is one of the following:
% 
\begin{indentedcodelist}
ref
val
uval
\end{indentedcodelist}
\end{fortranspecific}


\descr
The \code{linear} clause provides a superset of the functionality provided by the \code{private} clause.
A list item that appears in a \code{linear} clause is subject to the \code{private} clause semantics described
in \specref{subsubsec:private clause} except as noted.
If \plc{linear-step} is not specified, it is assumed to be 1.

When a \code{linear} clause is specified on a construct, the value of the new list item on each iteration of the associated loop(s) corresponds to the value of the original list item before entering the construct plus the logical number of the iteration times \plc{linear-step}.
The value corresponding to the sequentially last iteration of the associated loop(s) is assigned to the original list item.

When a \code{linear} clause is specified on a declarative directive, all list items must be formal parameters (or, in Fortran, dummy arguments) of a function that will be invoked concurrently on each SIMD lane.
If no \plc{modifier} is specified or the \code{val} or \code{uval} modifier is specified, the value of each list item on each lane corresponds to the value of the list item upon entry to the function plus the logical number of the lane times \plc{linear-step}.
If the \code{uval} modifier is specified, each invocation uses the same storage location for each SIMD lane; this storage location is updated with the final value of the logically last lane.
If the \code{ref} modifier is specified, the storage location of each list item on each lane corresponds to an array at the storage location upon entry to the function indexed by the logical number of the lane times \plc{linear-step}.


\restrictions
\begin{itemize}
\item The \plc{linear-step} expression must be invariant during the execution of the region
corresponding to the construct. Otherwise, the execution results in unspecified
behavior.

\item Only a loop iteration variable of a loop that is associated with the
    construct may appear as a \plc{list-item} in a \code{linear} clause if a
    \code{reduction} clause with the \code{inscan} modifier also appears on
    the construct.

\begin{cspecific}
\item A \plc{list-item} that appears in a \code{linear} clause must be of integral or pointer type.
\end{cspecific}

\begin{cppspecific}
\item A \plc{list-item} that appears in a \code{linear} clause without the \code{ref} modifier must be of integral or pointer type, or must be a reference to an integral or pointer type.
\item The \code{ref} or \code{uval} modifier can only be used if the \plc{list-item} is of a reference type.
\item If a list item in a \code{linear} clause on a worksharing
construct has a reference type then it must bind to the same object for all threads of the team.

\item If the list item is of a reference type and the \code{ref} modifier is not specified and if any write to the list item occurs before any read of the list item then the result is unspecified.
\end{cppspecific}

\begin{fortranspecific}
\item A \plc{list-item} that appears in a \code{linear} clause without the \code{ref} modifier must be of type \code{integer}.
\item The \code{ref} or \code{uval} modifier can only be used if the \plc{list-item} is a dummy argument without the \code{VALUE} attribute.
\item Variables that have the \code{POINTER} attribute and Cray pointers may not appear in a \code{linear} clause.
\item If the list item has the \code{ALLOCATABLE} attribute and the \code{ref} modifier is not specified, the allocation status of the list item in the sequentially last iteration must be allocated upon exit from that iteration.
\item If the \code{ref} modifier is specified, variables with the \code{ALLOCATABLE} attribute, assumed-shape arrays and polymorphic variables may not appear in the \code{linear} clause.
\item If the list item is a dummy argument without the \code{VALUE} attribute and the \code{ref} modifier is not specified and if any write to the list item occurs before any read of the list item then the result is unspecified.
\item A common block name cannot appear in a \code{linear} clause.
\end{fortranspecific}
\end{itemize}










\subsection{Reduction Clauses and Directives}
\index{reduction clauses}
\label{subsec:Reduction Clauses and Directives}
The reduction clauses are data-sharing attribute clauses that can be used to perform some forms of recurrence
calculations (involving mathematically associative and commutative operators)
in parallel.

Reduction clauses include reduction scoping clauses and reduction participating
clauses. Reduction scoping clauses define the region in which a reduction is
computed. Reduction participating clauses define the participants in the
reduction.

Reduction clauses specify a \plc{reduction-identifier} and one or more list
items.










\subsubsection{Properties Common To All Reduction Clauses}
\label{subsubsec:Properties Common To All Reduction Clauses}

\syntax
The syntax of a \plc{reduction-identifier} is defined as follows:
\begin{cspecific} % L1 vvvvv
A \plc{reduction-identifier} is either an \plc{identifier} or one of the following operators:
\code{+},
\code{-},
\code{*},
\code{&},
\code{|},
\code{^},
\code{&&} and
\code{||}
\end{cspecific} % L1 ^^^^^

\begin{cppspecific} % L1 vvvvv
A \plc{reduction-identifier} is either an \plc{id-expression} or one of the following operators:
\code{+},
\code{-},
\code{*},
\code{&},
\code{|},
\code{^},
\code{&&} and
\code{||}
\end{cppspecific} % L1 ^^^^^

\begin{fortranspecific}
A \plc{reduction-identifier} is either a base language identifier, or a user-defined operator,
or one of the following operators:
\code{+},
\code{-},
\code{*},
\code{.and.},
\code{.or.},
\code{.eqv.},
\code{.neqv.},
 or one of the following intrinsic procedure names:
\code{max},
\code{min},
\code{iand},
\code{ior},
\code{ieor}.
\end{fortranspecific}
%
%
\begin{ccppspecific} % L0 vvvvv
Table~\ref{tab:Implicitly Declared C/C++ Reduction Identifiers} lists each
\plc{reduction-identifier} that is implicitly declared at every scope for
arithmetic types and its semantic initializer value. The actual initializer
value is that value as expressed in the data type of the reduction list item.



% Table
\nolinenumbers
\needspace{\sbns}
\renewcommand{\arraystretch}{1.5}
%\begin{center}
\tablecaption{Implicitly Declared C/C++ \plc{reduction-identifiers}\label{tab:Implicitly Declared C/C++ Reduction Identifiers}}
%% \tablecaption*{}
\tablefirsthead{%
\hline
\textsf{\textbf{Identifier}} & \textsf{\textbf{Initializer}} & \textsf{\textbf{Combiner}}\\
\hline \\[-3ex]
}
\tablehead{%
\multicolumn{2}{l}{\small\slshape table continued from previous page}\\
\hline
\textsf{\textbf{Identifier}} & \textsf{\textbf{Initializer}} & \textsf{\textbf{Combiner}}\\
\hline \\[-3ex]
}
\tabletail{%
\hline\\[-4ex]
\multicolumn{2}{l}{\small\slshape table continued on next page}\\
}
\tablelasttail{\hline}
\begin{supertabular}{ p{0.1\textwidth} p{0.3\textwidth} p{0.5\textwidth}}
{\scode{+}} & {\scode{omp_priv = 0}} & {\scode{omp_out += omp_in}}\\
{\scode{*}} & {\scode{omp_priv = 1}} & {\scode{omp_out *= omp_in}}\\
{\scode{-}} & {\scode{omp_priv = 0}} & {\scode{omp_out += omp_in}}\\
{\scode{&}} & {\scode{omp_priv =  ~ 0}} & {\scode{omp_out &= omp_in}}\\
{\scode{|}} & {\scode{omp_priv = 0}} & {\scode{omp_out |= omp_in}}\\
{\scode{^}} & {\scode{omp_priv = 0}} & {\scode{omp_out ^}}{\scode{= omp_in}}\\
{\scode{&&}} & {\scode{omp_priv = 1}} & {\scode{omp_out = omp_in && omp_out}}\\
{\scode{||}} & {\scode{omp_priv = 0}} & {\scode{omp_out = omp_in || omp_out}}\\
{\scode{max}} & {\scode{omp_priv = }\splc{Least representable number in the reduction list item type}} & {\scode{omp_out = omp_in > omp_out ? omp_in : omp_out}}\\
{\scode{min}} & {\scode{omp_priv = }\splc{Largest representable number in the reduction list item type}} & {\scode{omp_out = omp_in < omp_out ? omp_in : omp_out}}\\
\end{supertabular}
\bigskip
\end{ccppspecific} % L0 ^^^^^
\linenumbers
\bigskip

\begin{fortranspecific}
Table~\ref{tab:Implicitly Declared Fortran Reduction Identifiers} lists each
\plc{reduction-identifier} that is implicitly declared for numeric and logical
types and its semantic initializer value. The actual initializer value is that
value as expressed in the data type of the reduction list item.

% Table
\nolinenumbers
\renewcommand{\arraystretch}{1.5}
\tablefirsthead{%
\hline
\textsf{\textbf{Identifier}} & \textsf{\textbf{Initializer}} & \textsf{\textbf{Combiner}}\\
\hline \\[-3ex]
}
\tablehead{%
\multicolumn{2}{l}{\small\slshape table continued from previous page}\\
\hline
\textsf{\textbf{Identifier}} & \textsf{\textbf{Initializer}} & \textsf{\textbf{Combiner}}\\
\hline \\[-3ex]
}
\tabletail{%
\hline\\[-4ex]
\multicolumn{2}{l}{\small\slshape table continued on next page}\\
}
\tablelasttail{\hline}
\tablecaption{Implicitly Declared Fortran \plc{reduction-identifiers}\label{tab:Implicitly Declared Fortran Reduction Identifiers}}
\begin{supertabular}{ p{0.1\textwidth} p{0.30\textwidth} p{0.5\textwidth}}
{\scode{+}} & {\scode{omp_priv = 0}} & {\scode{omp_out = omp_in + omp_out}}\\
{\scode{*}} & {\scode{omp_priv = 1}} & {\scode{omp_out = omp_in * omp_out}}\\
{\scode{-}} & {\scode{omp_priv = 0}} & {\scode{omp_out = omp_in + omp_out}}\\
{\scode{.and.}} & {\scode{omp_priv = .true.}} & {\scode{omp_out = omp_in .and. omp_out}}\\
{\scode{.or.}} & {\scode{omp_priv = .false.}} & {\scode{omp_out = omp_in .or. omp_out}}\\
{\scode{.eqv.}} & {\scode{omp_priv = .true.}} & {\scode{omp_out = omp_in .eqv. omp_out}}\\
{\scode{.neqv.}} & {\scode{omp_priv = .false.}} & {\scode{omp_out = omp_in .neqv. omp_out}}\\
{\scode{max}} & {\scode{omp_priv = }\splc{Least representable number in the reduction list item type}} & {\scode{omp_out = max(omp_in, omp_out)}}\\
{\scode{min}} & {\scode{omp_priv = }\splc{Largest representable number in the reduction list item type}} & {\scode{omp_out = min(omp_in, omp_out)}}\\
{\scode{iand}} & {\scode{omp_priv = }\splc{All bits on}} & {\scode{omp_out = iand(omp_in, omp_out)}}\\
{\scode{ior}} & {\scode{omp_priv = 0}} & {\scode{omp_out = ior(omp_in, omp_out)}}\\
{\scode{ieor}} & {\scode{omp_priv = 0}} & {\scode{omp_out = ieor(omp_in, omp_out)}}\\
\end{supertabular}
\bigskip
\end{fortranspecific}
\linenumbers



In the above tables, \code{omp_in} and \code{omp_out} correspond to two
identifiers that refer to storage of the type of the list item. \code{omp_out}
holds the final value of the combiner operation.

Any \plc{reduction-identifier} that is defined with the \code{declare}~\code{reduction}
directive is also valid. In that case, the initializer and combiner of the
\plc{reduction-identifier} are specified by the \plc{initializer-clause} and
the \plc{combiner} in the \code{declare}~\code{reduction} directive.




\descr
A reduction clause specifies a \plc{reduction-identifier} and one or more list
items.

The \plc{reduction-identifier} specified in a reduction clause must match a
previously declared \plc{reduction-identifier} of the same name and type for
each of the list items. This match is done by means of a name lookup in the
base language.

The list items that appear in a \code{reduction} clause may include array sections.


\begin{cppspecific}
If the type is a derived class, then any \plc{reduction-identifier} that
matches its base classes is also a match, if there is no specific match for the
type.

If the \plc{reduction-identifier} is not an \plc{id-expression}, then it is
implicitly converted to one by prepending the keyword operator (for example,
\code{+} becomes {\plc{operator}\code{+}}).

If the \plc{reduction-identifier} is qualified then a qualified name lookup is
used to find the declaration.

If the \plc{reduction-identifier} is unqualified then an \emph{argument-dependent name lookup}
must be performed using the type of each list item.
\end{cppspecific}

If the list item is an array or array section, it will be treated as
if a reduction clause would be applied to each separate element
of the array section.

Any copies associated with the reduction are initialized with the intializer
value of the \plc{reduction-identifier}.

Any copies are combined using the combiner associated with the
\plc{reduction-identifier}.


\restrictions
The restrictions common to reduction clauses are as follows:

\begin{itemize}
\item Any number of reduction clauses can be specified on the directive, but a
list item (or any array element in an array section) can appear only once in
reduction clauses for that directive.

\item For a \plc{reduction-identifier} declared with the \code{declare}~\code{reduction}
construct, the directive must appear before its use in a reduction clause.

\item If a list item is an array section or an array element, its base expression must be a base
language identifier.

\item If a list item is an array section, it must specify contiguous storage
and it cannot be a zero-length array section.

\item If a list item is an array section, accesses to the elements of the array
outside the specified array section result in unspecified behavior.

\begin{cspecific}
\item A variable that is part of another variable, with the exception of array
elements, cannot appear in a reduction clause.
\end{cspecific}

\begin{cppspecific}
\item A variable that is part of another variable, with the exception of array
elements, cannot appear in a reduction clause except if the reduction clause
is associated with a construct within a class non-static member function and
the variable is an accessible data member of the object for which the
non-static member function is invoked.
\end{cppspecific}

\begin{ccppspecific}
\item The type of a list item that appears in a reduction clause must be valid for the
\plc{reduction-identifier}. For a \code{max} or \code{min} reduction in C, the type of the list item must
be an allowed arithmetic data type: \code{char}, \code{int}, \code{float}, \code{double}, or \code{_Bool},
possibly modified with \code{long}, \code{short}, \code{signed}, or \code{unsigned}. For a \code{max} or \code{min}
reduction in C++, the type of the list item must be an allowed arithmetic data type:
\code{char}, \code{wchar_t}, \code{int}, \code{float}, \code{double}, or \code{bool}, possibly modified with \code{long},
\code{short}, \code{signed}, or \code{unsigned}.

\item A list item that appears in a reduction clause must not be \code{const}-qualified.

\item The \plc{reduction-identifier} for any list item must be unambiguous and accessible.
\end{ccppspecific}
\bigskip

\begin{fortranspecific}
\item A variable that is part of another variable, with the exception of array
elements, cannot appear in a reduction clause.

\item A type parameter inquiry cannot appear in a reduction clause.

\item The type, type parameters and rank of a list item that appears in a reduction clause
must be valid for the \plc{combiner} and \plc{initializer}.

\item A list item that appears in a reduction clause must be definable.

\item A procedure pointer may not appear in a reduction clause.

\item A pointer with the \code{INTENT(IN)} attribute may not appear in the
reduction clause.

\item An original list item with the \code{POINTER} attribute or any pointer
component of an original list item that is referenced in the \plc{combiner}
must be associated at entry to the construct that contains the reduction
clause. Additionally, the list item or the pointer component of the list item
must not be deallocated, allocated, or pointer assigned within the region.

\item An original list item with the \code{ALLOCATABLE} attribute or any
allocatable component of an original list item that corresponds to the special
variable identifier in the \plc{combiner} or the \plc{initializer} must be
in the allocated state at entry to the construct that contains the reduction
clause. Additionally, the list item or the allocatable component of the list
item must be neither deallocated nor allocated, explicitly or implicitly,
within the region.

\item If the \plc{reduction-identifier} is defined in a \code{declare}~\code{reduction}
directive, the \code{declare}~\code{reduction} directive must be in the same
subprogram, or accessible by host or use association.

\item If the \plc{reduction-identifier} is a user-defined operator, the same
explicit interface for that operator must be accessible as at the
\code{declare}~\code{reduction} directive.

\item If the \plc{reduction-identifier} is defined in a \code{declare}~\code{reduction}
directive, any subroutine or function referenced in the initializer clause or
combiner expression must be an intrinsic function, or must have an explicit
interface where the same explicit interface is accessible as at the
\code{declare}~\code{reduction} directive.
\end{fortranspecific}
\end{itemize}

\events

The \plc{reduction-begin} event occurs before a task 
begins to perform loads and stores that belong to the implementation of a reduction
and the \plc{reduction-end} event occurs after the task 
has completed loads and stores associated with the reduction.
If a task participates in multiple reductions, each reduction may be bracketed by its own pair of 
\plc{reduction-begin}/\plc{reduction-end} events or multiple reductions may be bracketed by a single pair of events.
The interval defined by a pair of \plc{reduction-begin}/\plc{reduction-end} events may not contain a task scheduling point.


\tools

A thread dispatches a registered \code{ompt_callback_reduction}
for each occurrence of a \plc{reduction-begin} or \plc{reduction-end} 
event in that thread.  The callback occurs in the context of the task performing the reduction.
This callback has the type signature \code{ompt_callback_sync_region_t}.  
The callback receives \code{ompt_sync_region_reduction} in its \plc{kind} argument and
\code{ompt_scope_begin} or \code{ompt_scope_end} as its
\plc{endpoint} argument, as appropriate.

\crossreferences
\begin{itemize}

\item \code{ompt_scope_begin} and \code{ompt_scope_end}, see
  \specref{sec:ompt_scope_endpoint_t}.

\item \code{ompt_sync_region_reduction}, see
  \specref{sec:ompt_sync_region_t}.

\item \code{ompt_callback_sync_region_t}, see
  \specref{sec:ompt_callback_sync_region_t}.

\end{itemize}


\subsubsection{Reduction Scoping Clauses}
\label{subsubsec:Reduction Scoping Clauses}
Reduction scoping clauses define the region in which a reduction is computed by
tasks or SIMD lanes. All properties common to all reduction clauses,
which are defined in Section~\ref{subsubsec:Properties Common To All Reduction
Clauses}, apply to reduction scoping clauses.

The number of copies created for each list item and the time at which those
copies are initialized are determined by the particular reduction scoping clause
that appears on the construct.

The time at which the original list item contains the result of the reduction
is determined by the particular reduction scoping clause.

\begin{samepage}
\begin{fortranspecific}
If the original list item has the \code{POINTER} attribute, copies of
the list item are associated with private targets.
\end{fortranspecific}
\end{samepage}

If the list item is an array section, the elements of any copy of the array section will
be allocated contiguously.

The location in the OpenMP program at which values are combined and the
order in which values are combined are unspecified. Therefore, when
comparing sequential and parallel runs, or when comparing one parallel run to
another (even if the number of threads used is the same), there is no guarantee
that bit-identical results will be obtained or that side effects (such as
floating-point exceptions) will be identical or take place at the same location
in the OpenMP program.

To avoid race conditions, concurrent reads or updates of the original list item
must be synchronized with the update of the original list item that occurs as a
result of the reduction computation.










\subsubsection{Reduction Participating Clauses}
\label{subsubsec:Reduction Participating Clauses}
A reduction participating clause specifies a task or a SIMD lane as a
participant in a reduction defined by a reduction scoping clause.
All properties common to all reduction clauses, which are defined in
Section~\ref{subsubsec:Properties Common To All Reduction Clauses}, apply to
reduction participating clauses.

Accesses to the original list item may be replaced by accesses to copies of the
original list item created by a region corresponding to a construct with a
reduction scoping clause.

In any case, the final value of the reduction must be determined as if all tasks
or SIMD lanes that participate in the reduction are executed sequentially in
some arbitrary order.










\subsubsection{\hcode{reduction} Clause}
\index{reduction@{\code{reduction}}}
\index{clauses!reduction@{\code{reduction}}}
\label{subsubsec:reduction clause}
\summary
The \code{reduction} clause specifies a \plc{reduction-identifier} and one or
more list items. For each list item, a private copy is created in each implicit
task or SIMD lane and is initialized with the initializer value of the
\plc{reduction-identifier}. After the end of the region, the original list item
is updated with the values of the private copies using the combiner associated
with the \plc{reduction-identifier}.

\syntax
\begin{ompSyntax}
reduction(\plc{[ reduction-modifier},\plc{]reduction-identifier }:\plc{ list})
\end{ompSyntax}

Where \plc{reduction-identifier} is defined in Section
\ref{subsubsec:Properties Common To All Reduction Clauses}, and
\plc{reduction-modifier} is one of the following:
\begin{indentedcodelist}
inscan
task
default
\end{indentedcodelist}

\descr
The \code{reduction} clause is a reduction scoping clause and a reduction
participating clause, as described in Sections \ref{subsubsec:Reduction Scoping
Clauses} and \ref{subsubsec:Reduction Participating Clauses}. 

If \plc{reduction-modifier} is not present or the \code{default}
\plc{reduction-modifier} is present, the behavior is
as follows. For \code{parallel} and worksharing constructs, one or more
private copies of each list item are created for each implicit task, as if the
\code{private} clause had been used. For the \code{simd} construct, one or
more private copies of each list item are created for each SIMD lane, as if the
\code{private} clause had been used.  For the \code{taskloop} construct,
private copies are created according to the rules of the reduction scoping
clauses.  For the \code{teams} construct, one or more private copies of each
list item are created for the initial task of each team in the league, as if
the \code{private} clause had been used. For the \code{loop} construct,
private copies are created and used in the construct according to
Section~\ref{subsec:List Item Privatization}. At the end of a region
corresponding to an above construct for which the \code{reduction} clause was
specified, the original list item is updated by combining its original value
with the final value of each of the private copies, using the combiner of the
specified \plc{reduction-identifier}. 

If the \code{inscan} \plc{reduction-modifier} is present, a scan computation is
performed over updates to the list item performed in each logical iteration of
the loop associated with the worksharing-loop, worksharing-loop SIMD, or
\code{simd} construct (see Section~\ref{subsec:scan Directive}). The list
items are privatized in the construct according to Section~\ref{subsec:List
Item Privatization}.  At the end of the region, each original list item is
assigned the value of the private copy from the last logical iteration of the
loops associated with the construct.

If the \code{task} \plc{reduction-modifier} is present for a \code{parallel}
or worksharing construct, then each list item is privatized according to
Section~\ref{subsec:List Item Privatization}, and an unspecified number of
additional private copies are created to support task reductions.  Any copies
associated with the reduction are initialized before they are accessed by the
tasks participating in the reduction, which include all implicit tasks in the
corresponding region and all participating explicit tasks that specify an
\code{in_reduction} clause (see Section~\ref{subsubsec:in_reduction clause}).
After the end of the region, the original list item contains the result of the
reduction.

If \code{nowait} is not specified for the construct, the reduction computation will be complete at the end of the
construct; however, if the \code{reduction} clause is used on a construct to which \code{nowait} is
also applied, accesses to the original list item will create a race and, thus, have
unspecified effect unless synchronization ensures that they occur after all threads have
executed all of their iterations or \code{section} constructs, and the reduction computation
has completed and stored the computed value of that list item. This can most simply be
ensured through a barrier synchronization.

\restrictions
The restrictions to the \code{reduction} clause are as follows:

\begin{itemize}
\item All the common restrictions to all reduction clauses, which are listed in
Section \ref{subsubsec:Properties Common To All Reduction Clauses}, apply to
this clause.

\item A list item that appears in a \code{reduction} clause of a worksharing
construct must be shared in the \code{parallel} region to which a corresponding
worksharing region binds.

\item A list item that appears in a \code{reduction} clause with
the \code{inscan} \plc{reduction-modifier} must appear as a list item in an \code{inclusive} or
\code{exclusive} clause on a \code{scan} directive enclosed by the
construct.

\item A \code{reduction} clause without the \code{inscan} \plc{reduction-modifier} may not
appear on a construct on which a \code{reduction} clause with the
\code{inscan} \plc{reduction-modifier} appears.

\item A \code{reduction} clause with the \code{task} \plc{reduction-modifier}
may only appear on a \code{parallel} construct, a worksharing construct or a
combined or composite construct for which any of the aforementioned constructs is a
constituent construct and \code{simd} or \code{loop} are not constituent
constructs.

\item A \code{reduction} clause with the \code{inscan} \plc{reduction-modifier}
may only appear on a worksharing-loop construct, a worksharing-loop SIMD
construct, a \code{simd} construct, a parallel worksharing-loop construct or
a parallel worksharing-loop SIMD construct.

\item A list item that appears in a \code{reduction} clause of the innermost
enclosing worksharing or \code{parallel} construct may not be accessed in an
explicit task generated by a construct for which an \code{in_reduction} clause
over the same list item does not appear.

\item The \code{task} \plc{reduction-modifier} may not appear in a \code{reduction} clause if
the \code{nowait} clause is specified on the same construct.

\begin{ccppspecific}
\item If a list item in a \code{reduction} clause on a worksharing
construct has a reference type then it must bind to the same object for all threads of the team.

\item A variable of class type (or array thereof) that appears in a
\code{reduction} clause with the \code{inscan} \plc{reduction-modifier} requires an
accessible, umambiguous default constructor for the class type. The number of
calls to the default constructor while performing the scan computation is
unspecified.

\item A variable of class type (or array thereof) that appears in a
\code{reduction} clause with the \code{inscan} \plc{reduction-modifier} requires an
accessible, unambiguous copy assignment operator for the class type. The
number of calls to the copy assignment operator while
performing the scan computation is unspecified.
\end{ccppspecific}
\end{itemize}

\crossreferences
\begin{itemize}
    \item List Item Privatization, see \specref{subsec:List Item Privatization}.
    \item \code{private} clause, see \specref{subsubsec:private clause}.
    \item \code{scan} directive, see \specref{subsec:scan Directive}.
\end{itemize}







\subsubsection{\hcode{task_reduction} Clause}
\index{task_reduction@{\code{task_reduction}}}
\index{clauses!task_reduction@{\code{task_reduction}}}
\label{subsubsec:task_reduction clause}
\summary
The \code{task_reduction} clause specifies a reduction among tasks.

\syntax
\begin{ompSyntax}
task_reduction(\plc{reduction-identifier }:\plc{ list})
\end{ompSyntax}
Where \plc{reduction-identifier} is defined in Section
\ref{subsubsec:Properties Common To All Reduction Clauses}.

\descr
The \code{task_reduction} clause is a reduction scoping clause, as described in
\ref{subsubsec:Reduction Scoping Clauses}.

For each list item, the number of copies is unspecified. Any copies associated
with the reduction are initialized before they are accessed by the tasks
participating in the reduction. After the end of the region, the original list
item contains the result of the reduction.

\restrictions
The restrictions to the \code{task_reduction} clause are as follows:

\begin{itemize}
\item All the common restrictions to all reduction clauses, which are listed in
Section \ref{subsubsec:Properties Common To All Reduction Clauses}, apply to
this clause.
\end{itemize}












\subsubsection{\hcode{in_reduction} Clause}
\index{in_reduction@{\code{in_reduction}}}
\index{clauses!in_reduction@{\code{in_reduction}}}
\label{subsubsec:in_reduction clause}
\summary
The \code{in_reduction} clause specifies that a task participates in a reduction.

\syntax
\begin{ompSyntax}
in_reduction(\plc{reduction-identifier }:\plc{ list})
\end{ompSyntax}
Where \plc{reduction-identifier} is defined in Section \ref{subsubsec:Properties Common To All Reduction Clauses}

\descr

The \code{in_reduction} clause is a reduction participating clause, as
described in Section \ref{subsubsec:Reduction Participating Clauses}. For a
given a list item, the \code{in_reduction} clause defines a task to be a
participant in a task reduction that is defined by an enclosing region for a
matching list item that appears in a \code{task_reduction} clause or a
\code{reduction} clause with the \code{task} modifier, where either:
\begin{enumerate}

\item the matching list item has the same storage location as the list
    item in the \code{in_reduction} clause; or

\item a private copy, derived from the matching list item, that is used to
    perform the task reduction has the same storage location as the list item
    in the \code{in_reduction} clause.
\end{enumerate}

For the \code{task} construct, the generated task becomes the participating
task. For each list item, a private copy may be created as if the \code{private}
clause had been used.

For the \code{target} construct, the target task becomes the participating
task. For each list item, a private copy will be created in the data
environment of the target task as if the \code{private} clause had been used,
and this private copy will be implicitly mapped into the device data
environment of the target device.

At the end of the task region, if a private copy was created its value is
combined with a copy created by a reduction scoping clause or with the original
list item.

\restrictions
The restrictions to the \code{in_reduction} clause are as follows:

\begin{itemize}
\item All the common restrictions to all reduction clauses, which are listed in
Section \ref{subsubsec:Properties Common To All Reduction Clauses}, apply to
this clause.

\item For each list item, there must exist a matching list item that appears
    in a \code{task_reduction} clause or a \code{reduction} clause with the
    \code{task} modifier that is specified on a construct corresponding to a
    region in which the region of the participating task is closely nested.
    The construct corresponding to the innermost enclosing region that meets
    this condition must specify the same \plc{reduction-identifier} for the
    matching list item as the \code{in_reduction} clause.
\end{itemize}



\subsubsection{\hcode{declare}~\hcode{reduction} Directive}
\index{declare reduction@{\code{declare}~\code{reduction}}}
\index{directives!declare reduction@{\code{declare}~\code{reduction}}}
\label{subsubsec:declare reduction Directive}
\summary
The following section describes the directive for declaring user-defined reductions. The
\code{declare}~\code{reduction} directive declares a \plc{reduction-identifier} that can be used in a
\code{reduction} clause. The \code{declare}~\code{reduction} directive is a declarative directive.

\syntax
\begin{cspecific}
\begin{ompcPragma}
#pragma omp declare reduction(\plc{reduction-identifier }:\plc{ typename-list }:
\plc{combiner })\plc{[initializer-clause] new-line}
\end{ompcPragma}

where:

\begin{itemize}

\item \plc{reduction-identifier} is either a base language identifier or one of the following
operators:
\code{+},
\code{-},
\code{*},
\code{&},
\code{|},
\code{^},
\code{&&} and
\code{||}

\item \plc{typename-list} is a list of type names

\item \plc{combiner} is an expression

\item \plc{initializer-clause} is \code{initializer(}\plc{initializer-expr}\code{)}
where \plc{initializer-expr} is\linebreak
\code{omp_priv = }\plc{initializer} or \plc{function-name}\code{(}\plc{argument-list}\code{)}
\end{itemize}
\end{cspecific}


\begin{cppspecific}
\begin{ompcPragma}
#pragma omp declare reduction(\plc{reduction-identifier }:\plc{ typename-list }:
\plc{combiner})\plc{ [initializer-clause] new-line}
\end{ompcPragma}

where:

\begin{itemize}
\item \plc{reduction-identifier} is either an \plc{id-expression} or one of the following
operators:
\code{+},
\code{-},
\code{*},
\code{&},
\code{|},
\code{^},
\code{&&} and
\code{||}

\item \plc{typename-list} is a list of type names

\item \plc{combiner} is an expression

% An equal sign is intentionally missing for C++, so that
% initializer (omp_priv (4)) syntax is also valid in addition to
% initializer (omp_priv = 4).
\item \plc{initializer-clause} is \code{initializer(}\plc{initializer-expr}\code{)}
where \plc{initializer-expr} is\linebreak
\code{omp_priv} \plc{initializer} or \plc{function-name}\code{(}\plc{argument-list}\code{)}
\end{itemize}
\end{cppspecific}


\begin{fortranspecific}
\begin{ompfPragma}
!$omp declare reduction(\plc{reduction-identifier }:\plc{ type-list }:\plc{ combiner})
\plc{[initializer-clause]}
\end{ompfPragma}

where:

\begin{itemize}
\item \plc{reduction-identifier} is either a base language identifier, or a user-defined operator, or
one of the following operators:
\code{+},
\code{-},
\code{*},
\code{.and.},
\code{.or.},
\code{.eqv.},
\code{.neqv.}, or one of the following intrinsic procedure names:
\code{max},
\code{min},
\code{iand},
\code{ior},
\code{ieor}.

\item \plc{type-list} is a list of type specifiers that must not be \code{CLASS(*)} and abstract type

\item \plc{combiner} is either an assignment statement or a subroutine name followed by an
argument list

\item \plc{initializer-clause} is \code{initializer(}\plc{initializer-expr}\code{)},
where \plc{initializer-expr} is\linebreak
\code{omp_priv = }\plc{expression} or \plc{subroutine-name}\code{(}\plc{argument-list}\code{)}
\end{itemize}
\end{fortranspecific}

\descr
Custom reductions can be defined using the \code{declare}~\code{reduction} directive; the
\plc{reduction-identifier} and the type identify the \code{declare}~\code{reduction} directive. The
\plc{reduction-identifier} can later be used in a \code{reduction} clause using variables of the
type or types specified in the \code{declare}~\code{reduction} directive. If the directive applies
to several types then it is considered as if there were multiple \code{declare}~\code{reduction}
directives, one for each type.

\begin{fortranspecific}
If a type with deferred or assumed length type parameter is specified in a \code{declare}~\code{reduction} directive, the \plc{reduction-identifier} of that directive can be used in a \code{reduction} clause with any variable of the same type and the same kind parameter, regardless of the length type Fortran parameters with which the variable is declared.
\end{fortranspecific}

The visibility and accessibility of this declaration are the same as those of a variable
declared at the same point in the program. The enclosing context of the \plc{combiner} and of
the \plc{initializer-expr} will be that of the \code{declare}~\code{reduction} directive. The \plc{combiner}
and the \plc{initializer-expr} must be correct in the base language as if they were the body of
a function defined at the same point in the program.

\begin{fortranspecific}
If the \plc{reduction-identifier} is the same as the name of a user-defined operator or an extended operator, or the same as a generic name that is one of the allowed intrinsic procedures, and if the operator or procedure name appears in an accessibility statement in the same module, the accessibility of the corresponding \code{declare}~\code{reduction} directive is determined by the accessibility attribute of the statement.

If the \plc{reduction-identifier} is the same as a generic name that is one of the allowed intrinsic procedures and is accessible, and if it has the same name as a derived type in the same module, the accessibility of the corresponding \code{declare}~\code{reduction} directive is determined by the accessibility of the generic name according to the base language.
\end{fortranspecific}

\begin{cppspecific}
The \code{declare}~\code{reduction} directive can also appear at points in the program at which
a static data member could be declared. In this case, the visibility and accessibility of
the declaration are the same as those of a static data member declared at the same point
in the program.
\end{cppspecific}

The \plc{combiner} specifies how partial results can be combined into a single value. The
\plc{combiner} can use the special variable identifiers \code{omp_in} and \code{omp_out} that are of the
type of the variables being reduced with this \plc{reduction-identifier}. Each of them will
denote one of the values to be combined before executing the \plc{combiner}. It is assumed
that the special \code{omp_out} identifier will refer to the storage that holds the resulting
combined value after executing the \plc{combiner}.

The number of times the \plc{combiner} is executed, and the order of these executions, for
any \code{reduction} clause is unspecified.

\begin{fortranspecific}
If the \plc{combiner} is a subroutine name with an argument list, the \plc{combiner} is evaluated by
calling the subroutine with the specified argument list.

If the \plc{combiner} is an assignment statement, the \plc{combiner} is evaluated by executing the
assignment statement.
\end{fortranspecific}

As the \plc{initializer-expr} value of a user-defined reduction is not known \emph{a priori} the
\plc{initializer-clause} can be used to specify one. Then the contents of the \plc{initializer-clause}
will be used as the initializer for private copies of reduction list items where the
\code{omp_priv} identifier will refer to the storage to be initialized. The special identifier
\code{omp_orig} can also appear in the \plc{initializer-clause} and it will refer to the storage of the
original variable to be reduced.

The number of times that the \plc{initializer-expr} is evaluated, and the order of these
evaluations, is unspecified.

\begin{ccppspecific}
If the \plc{initializer-expr} is a function name with an argument list, the \plc{initializer-expr} is
evaluated by calling the function with the specified argument list. Otherwise, the
\plc{initializer-expr} specifies how \code{omp_priv} is declared and initialized.
\end{ccppspecific}
%\bigskip

\begin{cspecific}
If no \plc{initializer-clause} is specified, the private variables will be initialized following the
rules for initialization of objects with static storage duration.
\end{cspecific}

\begin{cppspecific}
If no \plc{initializer-expr} is specified, the private variables will be initialized following the
rules for \plc{default-initialization}.
\end{cppspecific}
%\bigskip

\begin{fortranspecific}
If the \plc{initializer-expr} is a subroutine name with an argument list, the \plc{initializer-expr} is
evaluated by calling the subroutine with the specified argument list.

If the \plc{initializer-expr} is an assignment statement, the \plc{initializer-expr} is evaluated by
executing the assignment statement.

If no \plc{initializer-clause} is specified, the private variables will be initialized as follows:
\begin{itemize}
\item For \code{complex}, \code{real}, or \code{integer} types, the value 0 will be used.
\item For \code{logical} types, the value \code{.false.} will be used.
\item For derived types for which default initialization is specified, default initialization
will be used.
\item Otherwise, not specifying an \plc{initializer-clause} results in unspecified behavior.
\end{itemize}
\end{fortranspecific}
%\bigskip

\begin{ccppspecific}
If \plc{reduction-identifier} is used in a \code{target} region then a \code{declare}~\code{target} construct
must be specified for any function that can be accessed through the \plc{combiner} and
\plc{initializer-expr}.
\end{ccppspecific}
%\bigskip

\begin{fortranspecific}
If \plc{reduction-identifier} is used in a \code{target} region then a \code{declare}~\code{target} construct
must be specified for any function or subroutine that can be accessed through the \plc{combiner}
and \plc{initializer-expr}.
\end{fortranspecific}

\restrictions
\begin{itemize}

\item The only variables allowed in the \plc{combiner} are
  \code{omp_in} and \code{omp_out}.

\item The only variables allowed in the \plc{initializer-clause} are
  \code{omp_priv} and \code{omp_orig}.

\item If the variable \code{omp_orig} is modified in the \plc{initializer-clause}, the behavior is
unspecified.

\item If execution of the \plc{combiner} or the \plc{initializer-expr} results in the execution of an
OpenMP construct or an OpenMP API call, then the behavior is unspecified.

\item A \plc{reduction-identifier} may not be re-declared in the current scope for the same type
or for a type that is compatible according to the base language rules.

\item At most one \plc{initializer-clause} can be specified.

\item The \plc{typename-list} must not declare new types.

\begin{ccppspecific}
\item A type name in a \code{declare}~\code{reduction} directive cannot be a function type, an
array type, a reference type, or a type qualified with \code{const}, \code{volatile} or
\code{restrict}.
\end{ccppspecific}
\bigskip

\begin{cspecific}
\item If the \plc{initializer-expr} is a function name with an argument list, then one of the
arguments must be the address of \code{omp_priv}.
\end{cspecific}
% \bigskip

\begin{cppspecific}
\item If the \plc{initializer-expr} is a function name with an argument list, then one of the
arguments must be \code{omp_priv} or the address of \code{omp_priv}.
\end{cppspecific}
% \bigskip

\begin{fortranspecific}
\item If the \plc{initializer-expr} is a subroutine name with an argument list, then one of the
arguments must be \code{omp_priv}.

\item If the \code{declare}~\code{reduction} directive appears in the specification part of a module and the corresponding reduction clause does not appear in the same module, the \plc{reduction-identifier} must be the same as the name of a user-defined operator, one of the allowed operators that is extended or a generic name that is the same as the name of one of the allowed intrinsic procedures.

\item If the \code{declare}~\code{reduction} directive appears in the specification of a module, if the corresponding \code{reduction} clause does not appear in the same module, and if the \plc{reduction-identifier} is the same as the name of a user-defined operator or an extended operator, or the same as a generic name that is the same as one of the allowed intrinsic procedures then the interface for that operator or the generic name must be defined in the specification of the same module, or must be accessible by use association.

\item Any subroutine or function used in the \code{initializer} clause or \plc{combiner} expression must be an intrinsic function, or must have an accessible interface.

\item Any user-defined operator, defined assignment or extended operator used in the \code{initializer} clause or \plc{combiner} expression must have an accessible interface.

\item If any subroutine, function, user-defined operator, defined assignment or extended operator is used in the \code{initializer} clause or \plc{combiner} expression, it must be accessible to the subprogram in which the corresponding \code{reduction} clause is specified.

\item If the length type parameter is specified for a type, it must be a constant, a colon or an~\code{*}.

\item If a type with deferred or assumed length parameter is specified in a \code{declare}~\code{reduction} directive, no other \code{declare}~\code{reduction} directive with the same type, the same kind parameters and the same \plc{reduction-identifier} is allowed in the same scope.

\item Any subroutine used in the \code{initializer} clause or \plc{combiner} expression must not have any alternate returns appear in the argument list.
\end{fortranspecific}
\end{itemize}

\crossreferences
\begin{itemize}
\item \code{reduction} clause,
\specref{subsubsec:reduction clause}.
\end{itemize}







\subsection{Data Copying Clauses}
\label{subsec:Data Copying Clauses}
\index{data copying clauses}
\index{clauses!data copying}
This section describes the \code{copyin} clause (allowed on the \code{parallel} directive and
combined parallel worksharing directives) and the \code{copyprivate} clause (allowed on
the \code{single} directive).

These clauses support the copying of data values from private or threadprivate variables
on one implicit task or thread to the corresponding variables on other implicit tasks or
threads in the team.

The clauses accept a comma-separated list of list items (see \specref{sec:Directive Format}).
All list items appearing in a clause must be visible, according to the scoping rules of the
base language. Clauses may be repeated as needed, but a list item that specifies a given
variable may not appear in more than one clause on the same directive.

\begin{fortranspecific}
An associate name preserves the association with the selector established at the \code{ASSOCIATE} statement. A list item that appears in a data copying clause may be a selector of an \code{ASSOCIATE} construct. If the construct association is established prior to a parallel region, the association between the associate name and the original list item will be retained in the region.
\end{fortranspecific}







\subsubsection{\hcode{copyin} Clause}
\index{copyin@{\code{copyin}}}
\index{clauses!copyin@{\code{copyin}}}
\label{subsubsec:copyin clause}
\summary
The \code{copyin} clause provides a mechanism to copy the value of the master thread's
threadprivate variable to the threadprivate variable of each other member of the team
executing the \code{parallel} region.

\syntax
The syntax of the \code{copyin} clause is as follows:

\begin{ompSyntax}
copyin(\plc{list})
\end{ompSyntax}

\descr
\begin{ccppspecific}
The copy is done after the team is formed and prior to the start of execution of the
associated structured block. For variables of non-array type, the copy occurs by copy
assignment. For an array of elements of non-array type, each element is copied as if by
assignment from an element of the master thread's array to the corresponding element of
the other thread's array.
\end{ccppspecific}

\begin{cppspecific}
For class types, the copy assignment operator is invoked. The order in which copy
assignment operators for different variables of class type are called is unspecified.
\end{cppspecific}

\begin{fortranspecific}
The copy is done, as if by assignment, after the team is formed and prior to the start of
execution of the associated structured block.

On entry to any \code{parallel} region, each thread's copy of a variable that is affected by
a \code{copyin} clause for the \code{parallel} region will acquire the
type parameters, allocation, association, and
definition status of the master thread's copy, according to the following rules:

\begin{itemize}
\item If the original list item has the \code{POINTER} attribute, each copy receives the same
association status of the master thread's copy as if by pointer assignment.

\item If the original list item does not have the \code{POINTER}
attribute, each copy becomes defined with the value of the master
thread's copy as if by intrinsic assignment unless the list item has
a type bound procedure as a defined assignment.  If the original
list item that does not have the \code{POINTER} attribute has the
allocation status of unallocated, each copy will have the same
status.

\item If the original list item is unallocated or unassociated, the
thread's copy of the variable inherits the declared type parameters
and the default type parameter values from the original list item.
\end{itemize}
\end{fortranspecific}

\restrictions
The restrictions to the \code{copyin} clause are as follows:
\begin{ccppspecific}
\begin{itemize}
\item A list item that appears in a \code{copyin} clause must be threadprivate.

\item A variable of class type (or array thereof) that appears in a \code{copyin} clause requires
an accessible, unambiguous copy assignment operator for the class type.
\end{itemize}
\end{ccppspecific}

\begin{fortranspecific}
\begin{itemize}
\item A list item that appears in a \code{copyin} clause must be threadprivate. Named variables
appearing in a threadprivate common block may be specified: it is not necessary to
specify the whole common block.

\item A common block name that appears in a \code{copyin} clause must be declared to be a
common block in the same scoping unit in which the \code{copyin} clause appears.

\item If the list item is a polymorphic variable with the \code{ALLOCATABLE} attribute, the behavior is unspecified.
\end{itemize}
\end{fortranspecific}









\subsubsection{\hcode{copyprivate} Clause}
\index{copyprivate@{\code{copyprivate}}}
\index{clauses!copyprivate@{\code{copyprivate}}}
\label{subsubsec:copyprivate clause}
\summary
The \code{copyprivate} clause provides a mechanism to use a private variable to broadcast
a value from the data environment of one implicit task to the data environments of the
other implicit tasks belonging to the \code{parallel} region.

To avoid race conditions, concurrent reads or updates of the list item must be
synchronized with the update of the list item that occurs as a result of the
\code{copyprivate} clause.

\syntax
The syntax of the \code{copyprivate} clause is as follows:

\begin{ompSyntax}
copyprivate(\plc{list})
\end{ompSyntax}

\descr
The effect of the \code{copyprivate} clause on the specified list items occurs after the
execution of the structured block associated with the \code{single} construct (see
\specref{subsec:single Construct}),
and before any of the threads in the team have left the barrier
at the end of the construct.

\begin{ccppspecific}
In all other implicit tasks belonging to the \code{parallel} region, each specified list item
becomes defined with the value of the corresponding list item in the implicit task associated with the
thread that executed the structured block. For variables of non-array type, the definition
occurs by copy assignment. For an array of elements of non-array type, each element is
copied by copy assignment from an element of the array in the data environment of the
implicit task associated with the thread that executed the structured block to the
corresponding element of the array in the data environment of the other implicit tasks
\end{ccppspecific}

\begin{cppspecific}
For class types, a copy assignment operator is invoked. The order in which copy
assignment operators for different variables of class type are called is unspecified.
\end{cppspecific}

\begin{fortranspecific}
If a list item does not have the \code{POINTER} attribute, then in all other implicit tasks
belonging to the \code{parallel} region, the list item becomes defined as if by intrinsic
assignment with the value of the corresponding list item in the implicit task associated
with the thread that executed the structured block.  If the list item
has a type bound procedure as a defined assignment, the assignment is
performed by the defined assignment.

If the list item has the \code{POINTER} attribute, then, in all other implicit tasks belonging to
the \code{parallel} region, the list item receives, as if by pointer assignment, the same
association status of the corresponding list item in the implicit task associated with the
thread that executed the structured block.

The order in which any final subroutines for different variables of a finalizable type are called is unspecified.
\end{fortranspecific}

\begin{note}
The \code{copyprivate} clause is an alternative to using a shared variable for the
value when providing such a shared variable would be difficult (for example, in a
recursion requiring a different variable at each level).
\end{note}

\restrictions
The restrictions to the \code{copyprivate} clause are as follows:

\begin{itemize}
\item All list items that appear in the \code{copyprivate} clause must be either threadprivate
or private in the enclosing context.

\item A list item that appears in a \code{copyprivate} clause may not appear in a \code{private} or
\code{firstprivate} clause on the \code{single} construct.

\begin{cppspecific}
\item A variable of class type (or array thereof) that appears in a \code{copyprivate} clause
requires an accessible unambiguous copy assignment operator for the class type.
\end{cppspecific}

\begin{fortranspecific}
\item A common block that appears in a \code{copyprivate} clause must be threadprivate.

\item Pointers with the \code{INTENT(IN)} attribute may not appear in the \code{copyprivate}
clause.
\item The list item with the \code{ALLOCATABLE} attribute must have the allocation status of allocated when the intrinsic assignment is performed.

\item If the list item is a polymorphic variable with the \code{ALLOCATABLE} attribute, the behavior is unspecified.
\end{fortranspecific}
\end{itemize}







%%% map

\subsection{Data-Mapping Attribute Rules, Clauses, and Directives}
\label{subsec:Data-Mapping Attribute Rules, Clauses, and Directives}
\index{data-mapping rules and clauses}
\index{attributes, data-mapping}

%% Do we need something about
%% ``explicit \code{declare}~\code{target} directives''?
This section describes how the data-mapping and data-sharing attributes of
any variable referenced in a \code{target} region are determined. When specified,
explicit data-sharing attributes, \code{map} or \code{is_device_ptr} clauses on
\code{target} directives determine these attributes.  Otherwise, the first matching rule
from the following implicit data-mapping rules applies for variables referenced
in a \code{target} construct that are not declared in the construct and do not appear in
data-sharing attribute, \code{map} or \code{is_device_ptr} clauses:

\begin{itemize}
\item If a variable appears in a \code{to} or \code{link} clause on a
  \code{declare}~\code{target} directive then it is treated as if it had
  appeared in a \code{map} clause with a \plc{map-type} of \code{tofrom}.

\item If a list item appears in a \code{reduction}, \code{lastprivate} or
    \code{linear} clause on a combined \code{target} construct then it is treated as
    if it also appears in a \code{map} clause with a \plc{map-type} of
    \code{tofrom}.

\item If a list item appears in an \code{in_reduction} clause on a
    \code{target} construct then it is treated as if it also appears in a
    \code{map} clause with a \plc{map-type} of \code{tofrom} and a
    \plc{map-type-modifier} of \code{always}.

\item If a \code{defaultmap} clause is present for the category of the
    variable and specifies an implicit behavior other than \code{default},
    the data-mapping attribute is determined by that clause.

\begin{cppspecific}

\item If the \code{target} construct is within a class non-static member
    function, and a variable is an accessible data member of the object for
    which the non-static data member function is invoked, the variable is
    treated as if the \code{this[:1]} expression had appeared in a \code{map}
    clause with a \plc{map-type} of \code{tofrom}.  Additionally, if the
    variable is of a type pointer or reference to pointer, it is also
    treated as if it has appeared in a \code{map} clause as a zero-length
    array section.

\item If the \code{this} keyword is referenced inside a \code{target}
    construct within a class non-static member function, it is treated as
    if the \code{this[:1]} expression had appeared in a \code{map} clause
    with a \plc{map-type} of \code{tofrom}.

\end{cppspecific}

\begin{ccppspecific}

\item A variable that is of type pointer is treated as if it is the base
    pointer of a zero-length array section that appeared as a list item in a
    \code{map} clause.  

\end{ccppspecific}

\begin{cppspecific}

\item A variable that is of type reference to pointer is treated as if it
    had appeared in a \code{map} clause as a zero-length array section.

\end{cppspecific}

\item If a variable is not a scalar then it is treated as if it had appeared 
    in a \code{map} clause with a \plc{map-type} of \code{tofrom}.

\begin{fortranspecific}
\item If a scalar variable has the \code{TARGET}, \code{ALLOCATABLE}
or \code{POINTER} attribute then
it is treated as if it has appeared in a \code{map} clause with a
\plc{map-type} of \code{tofrom}.
\end{fortranspecific}

\item If none of the above rules applies then a scalar variable is not
mapped, but instead has an implicit data-sharing attribute of
firstprivate (see \specref{subsubsec:Variables Referenced in a Construct}).
\end{itemize}

\subsubsection{\hcode{map} Clause}
\label{subsec:map Clause}
\index{map@{\code{map}}}
\index{clauses!map@{\code{map}}}
\summary
The \code{map} clause specifies how an original list item is mapped from the current task's data environment to a corresponding list item in the device data environment of the device identified by the construct.

\syntax
The syntax of the map clause is as follows:

\begin{ompSyntax}
map(\plc{[ [map-type-modifier[},\plc{] [map-type-modifier[},\plc{] ...] map-type} : \plc{] locator-list})
\end{ompSyntax}

where \plc{map-type} is one of the following:

\begin{indentedcodelist}
to
from
tofrom
alloc
release
delete
\end{indentedcodelist}

and \plc{map-type-modifier} is one of the following:

\begin{indentedcodelist}
always
close
mapper(\plc{mapper-identifier})
\end{indentedcodelist}

\descr
The list items that appear in a \code{map} clause may include array sections and structure elements.



The \plc{map-type} and \plc{map-type-modifier} specify the effect of the \code{map} clause, as described below.

For a given construct, the effect of a \code{map} clause with the \code{to},
\code{from}, or \code{tofrom} \plc{map-type} is ordered before the effect of a
\code{map} clause with the \code{alloc}, \code{release}, or \code{delete}
\plc{map-type}. If a mapper is specified for the type being mapped, or
explicitly specified with the \verb`mapper` \plc{map-type-modifier}, then the
effective \code{map-type} of a list item will be determined according to the
rules of map-type decay.

If a mapper is specified for the type being mapped, or explicitly specified with
the \verb`mapper` \plc{map-type-modifier}, then all map clauses that appear on the
\verb`declare mapper` directive are treated as though they appeared on the
construct with the \verb`map` clause.  Array sections of a mapper type are
mapped as normal, then each element in the array section is mapped according to
the rules of the mapper.

\begin{ccppspecific}
If a list item in a \code{map} clause is a variable of structure type then it
is treated as if each structure element contained in the variable is a list
item in the clause.
\end{ccppspecific}

\begin{fortranspecific}
If a list item in a \code{map} clause is a derived type variable
then it is treated as if each nonpointer component is a list item in
the clause.
\end{fortranspecific}

If a list item in a \code{map} clause is a structure element then all other
structure elements (except pointer component, for Fortran) of the
containing structure variable form a \plc{structure sibling list}.  The \code{map} clause and the structure sibling list are
associated with the same construct.  If a corresponding list item of the
structure sibling list item is present in the device data environment when the
construct is encountered then:
    \begin{itemize}

    \item If the structure sibling list item does not appear in a \code{map}
    clause on the construct then:
    \begin{itemize}

        \item If the construct is a \code{target}, \code{target}~\code{data},
        or \code{target}~\code{enter}~\code{data} construct then the structure
        sibling list item is treated as if it is a list item in a \code{map}
        clause on the construct with a \plc{map-type} of \code{alloc}.

        \item If the construct is \code{target}~\code{exit}~\code{data}
        construct, then the structure sibling list item is treated as if it is
        a list item in a \code{map} clause on the construct with a
        \plc{map-type} of \code{release}.

    \end{itemize}

    \item If the \code{map} clause in which the structure element appears as a
    list item has a \plc{map-type} of \code{delete} and the structure sibling
    list item does not appear as a list item in a \code{map} clause on the
    construct with a \plc{map-type} of \code{delete} then the structure sibling
    list item is treated as if it is a list item in a \code{map} clause on the
    construct with a \plc{map-type} of \code{delete}.

\end{itemize}

%If a structure is implicitly mapped and any part of the structure is
%explicitly mapped, only the corresponding storage for those explicitly
%mapped parts are accessible in the device data environment.
%
%If a structure that is not present in the device data environment is
%implicitly mapped by a construct and any part of the structure is
%explicitly mapped by the same construct, only the corresponding
%storage for those explicitly mapped parts are accessible in the device
%data environment as a result of the mapping of the structure.
%
If a structure or array that is not present in the device data
environment is implicitly mapped by a construct and any part of the
structure or array is explicitly mapped by the same construct, only
the corresponding storage for those explicitly mapped parts are
accessible in the device data environment as a result of the mapping
of the structure or array.

If \plc{item}\textsubscript{1} is a list item in a \code{map} clause, and
\plc{item}\textsubscript{2} is another list item in a \code{map} clause on the
same construct that has a base pointer that is, or is part of,
\plc{item}\textsubscript{1}, then:

\begin{itemize}

\item If the \code{map} clause(s) appear on a \code{target},
\code{target}~\code{data}, or \code{target}~\code{enter}~\code{data} construct,
then on entry to the corresponding region the effect of the \code{map} clause
on \plc{item}\textsubscript{1} is ordered to occur before the effect of the \code{map} clause
on \plc{item}\textsubscript{2}.

\item If the \code{map} clause(s) appear on a \code{target},
\code{target}~\code{data}, or \code{target}~\code{exit}~\code{data} construct,
then on exit from the corresponding region the effect of the \code{map} clause
on \plc{item}\textsubscript{2} is ordered to occur before the effect of the \code{map} clause
on \plc{item}\textsubscript{1}.

\end{itemize}

\begin{ccppspecific}
If a list item in a \code{map} clause has a base pointer, and a pointer
variable is present in the device data environment that corresponds to the
base pointer when the effect of the \code{map} clause occurs, then if the corresponding
list item is created in the device data environment: 

\begin{enumerate}

\item The corresponding pointer variable is assigned an address such that the
corresponding list item can be accessed through the pointer in a
\code{target} region.

\item The corresponding pointer variable becomes an attached pointer
for the corresponding list item.

\item If the original base pointer and the corresponding attached pointer
share storage, then the original list item and the corresponding list item 
must share storage. 

\end{enumerate}

\end{ccppspecific}

\begin{fortranspecific}
%If a list item is a derived type component that is not a pointer and
%it has a base pointer, the base pointer is treated as if it is a list
%item in the clause.
If a list item is not a pointer and it has a base pointer, the base
pointer is treated as if it is a list item in the clause.

If a list item has the \code{POINTER} attribute and if the association
status of the list item is associated, then it is treated as if the
pointer target is a list item in the clause.
%The corresponding pointer is associated with the corresponding pointer
%target in the device data environment.  The corresponding pointer
%becomes an attached pointer unless it was associated with the
%corresponding pointer target prior to a task encountering the
%construct associated with the \code{map} clause.
The corresponding pointer becomes associated with the corresponding
pointer target in the device data environment if either the
corresponding pointer or the corresponding pointer target is created
by a \code{map} clause on the construct.  The associated corresponding
pointer becomes an attached pointer unless it was associated with the
corresponding pointer target prior to a task encountering the
construct associated with the \code{map} clause.

\end{fortranspecific}

\begin{cppspecific}
If a \plc{lambda} is mapped explicitly or implicitly, variables
that are captured by the \plc{lambda} behave as follows:
\begin{itemize}
\item the variables that are of pointer type are treated as if they had
appeared in a \code{map} clause as zero-length array sections

\item the variables that are of reference type are treated as if they
had appeared in a \code{map} clause.
\end{itemize}

If a member variable is captured by a \plc{lambda} in class scope, and
the \plc{lambda} is later mapped explicitly or implicitly with its
full static type, the \plc{this} pointer is treated as if it had
appeared on a \code{map} clause.

\end{cppspecific}

The original and corresponding list items may share storage such that writes to either
item by one task followed by a read or write of the other item by another task without
intervening synchronization can result in data races.

If the \code{map} clause appears on a \code{target}, \code{target}~\code{data}, or \code{target}~\code{enter}~\code{data} construct then on entry to the region the following sequence of steps occurs as if performed as a single atomic operation:

% allocate storage
\begin{enumerate}
\item If a corresponding list item of the original list item is not present in the device data environment, then:
\begin{enumerate}
\item A new list item with language-specific attributes is derived from the original list item and created in the device data environment.
\item The new list item becomes the corresponding list item to the original list item in the device data environment.
\item The corresponding list item has a reference count that is initialized to zero.
\item The value of the corresponding list item is undefined.
\end{enumerate}

% incremenent the reference counter
\item If the corresponding list item's reference count was not already
incremented because of the effect of a \code{map} clause on the
construct then:
\begin{enumerate}
\item The corresponding list item's reference count is incremented by one
\end{enumerate}

% assign the corresponding variable
\item If the corresponding list item's reference count is one or the \code{always} \plc{map-type-modifier} is present, then:
\begin{enumerate}
\item If the \plc{map-type} is \code{to} or \code{tofrom}, then:
    \begin{itemize}

    \item For each part of the list item that is an attached pointer:  \begin{itemize}

        \item That part of the corresponding list item will
        have the value it had immediately prior to the effect of the \code{map} clause;

   \end{itemize}

    \item For each part of the list item that is not an attached pointer: \begin{itemize}

        \item The value of that part of the
        original list item is assigned to that part of the corresponding list item.

   \end{itemize}

    \item Concurrent reads or updates of any part
        of the corresponding list item must be synchronized with the update of the
        corresponding list item that occurs as a result of the \code{map} clause.

\end{itemize}
\end{enumerate}

\end{enumerate}

\begin{note}
If the effect of the \code{map} clauses on a construct would assign the
value of an original list item to a corresponding list item more than once,
then an implementation is allowed to ignore additional assignments of
the same value to the corresponding list item.
\end{note}

If the \code{map} clause appears on a \code{target}, \code{target}~\code{data}, or \code{target}~\code{exit}~\code{data} construct then on exit from the region the following sequence of steps occurs as if performed as a single atomic operation:
\begin{enumerate}
\item If a corresponding list item of the original list item is not present in the device data environment, then the list item is ignored.
\item If a corresponding list item of the original list item is present in the device data environment, then:
\begin{enumerate}

% decrement the ref count
\item If the corresponding list item's reference count is finite, then:
\begin{enumerate}
\item If the corresponding list item's reference count was not already
decremented because of the effect of a \code{map} clause on the
construct then:
\begin{enumerate}
\item If the \plc{map-type} is not \code{delete}, then the corresponding list
                item's reference count is decremented by one.
\end{enumerate}
\item If the \plc{map-type} is \code{delete}, then the corresponding list item's reference count is set to zero.
\end{enumerate}

% assign the original variable
\item If the corresponding list item's reference count is zero or the \code{always} \plc{map-type-modifier} is present, then:
\begin{enumerate}
\item If the \plc{map-type} is \code{from} or \code{tofrom} then:
\begin{itemize}

              \item For each part of the list item that is an attached pointer:  \begin{itemize}

                  \item That part of the original list item will
                    have the value it had immediately prior to the effect of the \code{map} clause;

                \end{itemize}

              \item For each part of the list item that is not an attached pointer: \begin{itemize}

                  \item The value of that part of the
                    corresponding list item is assigned to that part of the original list item;

                \end{itemize}

              \item To avoid race conditions: \begin{itemize}

                  \item Concurrent reads or updates of any part
                    of the original list item must be synchronized with the update of the
                    original list item that occurs as a result of the \code{map} clause;

                \end{itemize}


\end{itemize}
\end{enumerate}

% release the storage
\item If the corresponding list item's reference count is zero, then the corresponding list item is removed from the device data environment
\end{enumerate}
\end{enumerate}

\begin{note}
If the effect of the \code{map} clauses on a construct would assign the
value of a corresponding list item to an original list item more than once,
then an implementation is allowed to ignore additional assignments of
the same value to the original list item.
\end{note}

If a single contiguous part of the original storage of a list item with an
implicit data-mapping attribute has corresponding storage in the device data
environment prior to a task encountering the construct associated with the
\code{map} clause, only that part of the original storage will have
corresponding storage in the device data environment as a result of the \code{map}
clause.

\begin{ccppspecific}
If a new list item is created then a new list item of the same type, with automatic storage
duration, is allocated for the construct. The size and alignment of the new list
item are determined by the static type of the variable. This allocation occurs if the region
references the list item in any statement.
\end{ccppspecific}

\begin{fortranspecific}
If a new list item is created then a new list item of the same type, type parameter, and
rank is allocated.  The new list item inherits all default values for
the type parameters from the original list item.

If the allocation status of the original list item with the
\code{ALLOCATABLE} attribute is changed in the host device data
environment and the corresponding list item is already present in the
device data environment, the allocation status of the corresponding
list item is unspecified until a mapping operation is performed with a
\code{map} clause on entry to a \code{target}, \code{target}~\code{data}, or
\code{target}~\code{enter}~\code{data} region.
\end{fortranspecific}

The \plc{map-type} determines how the new list item is initialized.

If a \plc{map-type} is not specified, the \plc{map-type} defaults to \code{tofrom}.

The \code{close} \plc{map-type-modifier} is a hint to the runtime to allocate memory close to the target device.

\events
The \plc{target-map} event occurs when a thread maps data to or from a target device.

The \plc{target-data-op} event occurs when a thread initiates a data operation on a target device.

\tools

A thread dispatches a registered \code{ompt_callback_target_map}
callback for each occurrence of a \plc{target-map} event in that thread.
The callback occurs in the context of the target task.  The callback has type signature
\code{ompt_callback_target_map_t}.

A thread dispatches a registered \code{ompt_callback_target_data_op}
callback for each occurrence of a \plc{target-data-op} event in that thread.
The callback occurs in the context of the target task.  The callback has type signature
\code{ompt_callback_target_data_op_t}.

\restrictions
\begin{itemize}

\item A list item cannot appear in both a \code{map} clause and a data-sharing
    attribute clause on the same construct, unless the the construct is a
    combined construct.

\item Each of the \plc{map-type-modifier} modifiers can appear at most once on the
   \code{map} clause.

\begin{ccppspecific}

\item List items of the \code{map} clauses on the same construct must not share
original storage unless they are the same lvalue expression or array section.
\end{ccppspecific}

\item If a list item is an array section, it must specify contiguous storage.

\item If multiple list items are explicitly mapped on the same
construct and have the same containing array or have base pointers that share
original storage, and if any of the list items do not have corresponding list
items that are present in the device data environment prior to a task
encountering the construct, then the list items must refer to the same array
elements of either the containing array or the implicit array of the base
pointers.

\item If any part of the original storage of a list item with an
explicit data-mapping attribute has corresponding storage in the device data
environment prior to a task encountering the construct associated with the
\code{map} clause, all of the original storage must have corresponding storage
in the device data environment prior to the task encountering the construct.

\item If a list item is an element of a structure, and a different element
of the structure has a corresponding list item in the device data environment
prior to a task encountering the construct associated with the \code{map}
clause, then the list item must also have a corresponding list item in the
device data environment prior to the task encountering the construct.

\item A list item must have a mappable type.

\item \code{threadprivate} variables cannot appear in a \code{map} clause.

\item If a \code{mapper} map-type-modifier is specified, its type must match the type of the list-items passed to that map clause.

\item Memory spaces and memory allocators cannot appear as a list item in a \code{map} clause.

\begin{cppspecific}

\item If the type of a list item is a reference to a type \plc{T} then the
  reference in the device data environment is initialized to refer to the object
  in the device data environment that corresponds to the object referenced by
  the list item.  If mapping occurs, it occurs as though the object were mapped
  through a pointer with an array section of type \plc{T} and length one.

\item No type mapped through a reference can contain a reference to its own
  type, or any cycle of references to types that could produce a cycle of
  references.

\item If the list item is a \plc{lambda}, any pointers and references captured by
the \plc{lambda} must have the corresponding list item in the device
data environment prior to the task encountering the construct.
\end{cppspecific}

\begin{ccppspecific}
\item Initialization and assignment are through bitwise copy.
\item A list item cannot be a variable that is a member of a structure with a union type.

\item A bit-field cannot appear in a \code{map} clause.

\item A pointer that has a corresponding attached pointer may not be modified
      for the duration of the lifetime of the list item to which the
      corresponding pointer is attached in the device data environment.
\end{ccppspecific}
\end{itemize}

\begin{fortranspecific}
\begin{itemize}
\item List items of the \code{map} clauses on the same construct must not share
original storage unless they are the same variable or array section.

\item The value of the new list item becomes that of the original list item in the map
initialization and assignment.

\item If the allocation status of a list item or any subobject of the
  list item with the \code{ALLOCATABLE} attribute is unallocated upon
  entry to a \code{target} region, the list item or any subobject of
  the corresponding list item must be unallocated upon exit from the
  region.

\item If the allocation status of a list item or any subobject of the
  list item with the \code{ALLOCATABLE} attribute is allocated upon
  entry to a \code{target} region, the allocation status of the
  corresponding list item or any subobject of the corresponding list
  item must not be changed and must not be reshaped in the region.

\item If an array section is mapped and the size of the section is
  smaller than that of the whole array, the behavior of referencing
  the whole array in the \code{target} region is unspecified.

\item A list item must not be a whole array of an assumed-size array.

\item If the association status of a list item with the \code{POINTER}
  attribute is associated upon entry to a \code{target} region, the
  list item remains associated with the same pointer target upon exit
  from the region.

\item If the association status of a list item with the \code{POINTER}
  attribute is disassociated upon entry to a \code{target} region, the
  list item must be disassociated upon exit from the region.

\item If the association status of a list item with the \code{POINTER}
  attribute is undefined upon entry to a \code{target} region, the
  list item must be undefined upon exit from the region.

\item If the association status of a list item with the \code{POINTER}
  attribute is disassociated or undefined on entry and if the list
  item is associated with a pointer target inside a \code{target}
  region, then the pointer association status must become
  disassociated before the end of the region; otherwise the behavior
  is unspecified.

\end{itemize}
\end{fortranspecific}

\crossreferences
\begin{itemize}

\item \code{ompt_callback_target_map_t}, see
\specref{sec:ompt_callback_target_map_t}.

\item \code{ompt_callback_target_data_op_t}, see
\specref{sec:ompt_callback_target_data_op_t}.
\end{itemize}




\subsubsection{\hcode{defaultmap} Clause}
\label{subsubsec:defaultmap clause}
\index{defaultmap@{\code{defaultmap}}}
\index{clauses!defaultmap@{\code{defaultmap}}}

\summary

The \code{defaultmap} clause redefines the implicit data-mapping attributes of
variables that are referenced in a \code{target} construct and are implicitly
determined.

\syntax
The syntax of the \code{defaultmap} clause is as follows:

\begin{ompSyntax}
  defaultmap(\plc{implicit-behavior[:variable-category]})
\end{ompSyntax}

Where \plc{implicit-behavior} is one of:
\begin{indentedcodelist}
  alloc
  to
  from
  tofrom
  firstprivate
  none
  default
\end{indentedcodelist}

\begin{ccppspecific}
and \plc{variable-category} is one of:
\begin{indentedcodelist}
  scalar
  aggregate
  pointer
\end{indentedcodelist}
\end{ccppspecific}

\begin{fortranspecific}
and \plc{variable-category} is one of:
\begin{indentedcodelist}
  scalar
  aggregate
  allocatable
  pointer
\end{indentedcodelist}
\end{fortranspecific}

\descr
The \code{defaultmap} clause sets the implicit data-mapping attribute for all variables
referenced in the construct. If \plc{variable-category} is specified, the effect of
the \code{defaultmap} clause is as follows:

\begin{itemize}

  \item If \plc{variable-category} is \code{scalar}, all scalar variables of
    non-pointer type or all non-pointer non-allocatable scalar variables that
    have an implicitly determined data-mapping or data-sharing attribute will
    have a data-mapping or data-sharing attribute specified by
    \plc{implicit-behavior}.

  \item  If \plc{variable-category} is \code{aggregate} or \code{allocatable},
    all aggregate or allocatable variables that have an implicitly determined
    data-mapping or data-sharing attribute will have a data-mapping or
    data-sharing attribute specified by \plc{implicit-behavior}.

  \item If \plc{variable-category} is \code{pointer}, all variables of pointer
    type or with the POINTER attribute that have implicitly determined
    data-mapping or data-sharing attributes will have a data-mapping or
    data-sharing attribute specified by \plc{implicit-behavior}.  The
    zero-length array section and attachment an implicitly mapped pointer
    normally gets is only provided for the \code{default} behavior.

\end{itemize}

If no \plc{variable-category} is specified in the clause then
\plc{implicit-behavior} specifies the implicitly determined data-mapping or
data-sharing attribute for all variables referenced in the construct.  If
\plc{implicit-behavior} is \code{none}, each variable referenced in the
construct that does not have a predetermined data-sharing attribute and does
not appear in a \code{to} or \code{link} clause on a \code{declare}~\code{target}
directive must be listed in a data-mapping attribute clause, a data-sharing
attribute clause (including data-sharing attribute clause on a combined
construct where \code{target} is one of the constituent constructs), or
an \code{is_device_ptr} clause. If \plc{implicit-behavior} is \code{default},
then the clause has no effect for the variables in the category specified by
\plc{variable-category}.




\subsubsection{\hcode{declare}~\hcode{mapper} Directive}
\index{declare mapper@{\code{declare}~\code{mapper}}}
\index{directives!declare mapper@{\code{declare}~\code{mapper}}}
\index{constructs!declare mapper@{\code{declare}~\code{mapper}}}
\index{device constructs!declare mapper@{\code{declare}~\code{mapper}}}
\label{subsubsec:declare mapper Directive}

\summary

The \code{declare mapper} directive declares a user-defined mapper for a given
type, and may define a \plc{mapper-identifier} that can be used in a \code{map}
clause. The \code{declare mapper} directive is a declarative directive.

\syntax
\begin{ccppspecific}
The syntax of the \code{declare mapper} directive is as follows:

\begin{ompcPragma}
#pragma omp declare mapper(\plc{[mapper-identifier}:\plc{]type var})
\plc{[clause[ [},\plc{] clause] ... ] new-line}
\end{ompcPragma}

\end{ccppspecific}

\begin{fortranspecific}
The syntax of the \code{declare mapper} directive is as follows:

\begin{ompfPragma}
!$omp declare mapper(\plc{[mapper-identifier}:\plc{] type :: var})
\plc{[clause[ [},\plc{] clause] ... ]}
\end{ompfPragma}

\end{fortranspecific}

where:

\begin{itemize}
  \item \plc{mapper-identifier} is a base-language identifier or \pcode{default}
  \item \plc{type} is a valid type in scope (in Fortran, it must not be an abstract type)
  \item \plc{var} is a valid base-language identifier
  \item \plc{clause} is \pcode{map(}\plc{[[map-type-modifier[}\pcode{,}\plc{] [map-type-modifier[}\pcode{,}\plc{] ...]] map-type}\pcode{:}\plc{ ] list}\pcode{)}
  \item \plc{map-type} is one of the following:
    \begin{itemize}
      \item \code{alloc}
      \item \code{to}
      \item \code{from}
      \item \code{tofrom}
    \end{itemize}
  \item and \plc{map-type-modifier} is one of the following:
    \begin{itemize}
      \item \code{always}
      \item \code{close}
    \end{itemize}
\end{itemize}

\descr


User-defined mappers can be defined using the \code{declare mapper} directive.
The type and the \plc{mapper-identifier} uniquely identify the mapper for use in
a \code{map} clause later in the program.  If the \plc{mapper-identifier} is not
specified, then \code{default} is used.  The visibility and accessibility of
this declaration are the same as those of a variable declared at the same point
in the program.

The variable declared by \plc{var} is available for use in all \code{map}
clauses on the directive, and no part of the variable to be mapped is mapped by
default.

The default mapper for all types \plc{T}, designated by the pre-defined
\plc{mapper-identifier} \code{default}, is as follows unless a user-defined mapper is
specified for that type.

\begin{ompSyntax}

declare mapper(\plc{T} v) map(tofrom: v)

\end{ompSyntax}

Using the \code{default} \plc{mapper-identifier} overrides the pre-defined
default mapper for the given type, making it the default for all
variables of \plc{type}. All \code{map} clauses with this construct in scope
that map a list item of \plc{type} will use this mapper unless another is
explicitly specified.

All \code{map} clauses on the directive are expanded into corresponding
\code{map} clauses wherever this mapper is invoked, either by matching type or
by being explicitly named in a \code{map} clause. A \code{map} clause with list item
\plc{var} maps var as though no mapper were specified.


\begin{cppspecific}
The \code{declare mapper} directive can also appear at points in the program at
which a static data member could be declared. In this case, the visibility and
accessibility of the declaration are the same as those of a static data member
declared at the same point in the program.

\end{cppspecific}

\restrictions

\begin{itemize}

\item No instance of the mapper type can be mapped as part of the mapper, either
  directly or indirectly through another type, except the instance passed as the
  list item. If a set of \code{declare}~\code{mapper} directives results in a
  cyclic definition then the behavior is unspecified.

\item The \plc{type} must be of struct, union or class type in C and C++ or a
  non-intrisic type in Fortran.

\item The \plc{type} must not declare a new type.

\item At least one \code{map} clause that maps \plc{var} or at least one element
  of \plc{var} is required.

\item List-items in \code{map} clauses on this construct may only refer to the
  declared variable \plc{var} and entities that could be referenced by a
  procedure defined at the same location.

\item Each \plc{map-type-modifier} can appear at most once on the
   \code{map} clause.

 \item A \plc{mapper-identifier} may not be re-declared in the current scope for
   the same type or for a type that is compatible according to the base language
   rules.

\end{itemize}
