% This is an included file. See the master file for more information.
%
% When editing this file:
%
%    1. To change formatting, appearance, or style, please edit openmp.sty.
%
%    2. Custom commands and macros are defined in openmp.sty.
%
%    3. Be kind to other editors -- keep a consistent style by copying-and-pasting to
%       create new content.
%
%    4. We use semantic markup, e.g. (see openmp.sty for a full list):
%         \code{}     % for bold monospace keywords, code, operators, etc.
%         \plc{}      % for italic placeholder names, grammar, etc.
%
%    5. There are environments that provide special formatting, e.g. language bars.
%       Please use them whereever appropriate.  Examples are:
%
%         \begin{fortranspecific}
%         This is text that appears enclosed in blue language bars for Fortran.
%         \end{fortranspecific}
%
%         \begin{note}
%         This is a note.  The "Note -- " header appears automatically.
%         \end{note}
%
%    6. Other recommendations:
%         Use the convenience macros defined in openmp.sty for the minor headers
%         such as Comments, Syntax, etc.
%
%         To keep items together on the same page, prefer the use of
%         \begin{samepage}.... Avoid \parbox for text blocks as it interrupts line numbering.
%         When possible, avoid \filbreak, \pagebreak, \newpage, \clearpage unless that's
%         what you mean. Use \needspace{} cautiously for troublesome paragraphs.
%
%         Avoid absolute lengths and measures in this file; use relative units when possible.
%         Vertical space can be relative to \baselineskip or ex units. Horizontal space
%         can be relative to \linewidth or em units.
%
%         Prefer \emph{} to italicize terminology, e.g.:
%             This is a \emph{definition}, not a placeholder.
%             This is a \plc{var-name}.
%


\section{Worksharing Constructs}
\label{sec:Worksharing Constructs}
\index{worksharing constructs}
\index{constructs!worksharing}
\index{worksharing!constructs}
A worksharing construct distributes the execution of the corresponding region among 
the members of the team that encounters it. Threads execute portions of the region 
in the context of the implicit tasks that each one is executing. If the team consists 
of only one thread then the worksharing region is not executed in parallel.

A worksharing region has no barrier on entry; however, an implied barrier exists at 
the end of the worksharing region, unless a \code{nowait} clause is specified. If a 
\code{nowait} clause is present, an implementation may omit the barrier at the end 
of the worksharing region. In this case, threads that finish early may proceed 
straight to the instructions that follow the worksharing region without waiting for 
the other members of the team to finish the worksharing region, and without 
performing a flush operation.

The OpenMP API defines the worksharing constructs that are described in this section
as well as the worksharing-loop construct, which is described in 
\specref{subsec:Worksharing-Loop Construct}.

\begin{samepage}
\restrictions
The following restrictions apply to worksharing constructs:

\begin{itemize}
\item Each worksharing region must be encountered by all threads in a team or by none at
all, unless cancellation has been requested for the innermost enclosing parallel
region.

\item The sequence of worksharing regions and \code{barrier} regions encountered must be the
same for every thread in a team.
\end{itemize}
\end{samepage}



\subsection{\hcode{sections} Construct}
\label{subsec:sections Construct}
\index{sections@{\code{sections}}}
\index{constructs!sections@{\code{sections}}}
\summary
The \code{sections} construct is a non-iterative worksharing construct that 
contains a set of structured blocks that are to be distributed among and 
executed by the threads in a team. Each structured block is executed once by 
one of the threads in the team in the context of its implicit task.

\syntax
\begin{ccppspecific}
The syntax of the \code{sections} construct is as follows:

\begin{ompcPragma}
#pragma omp sections \plc{[clause[ [},\plc{] clause] ... ] new-line}
   {
   \plc{[}#pragma omp section \plc{new-line}\plc{]}
      \plc{structured-block}
   \plc{[}#pragma omp section \plc{new-line}
      \plc{structured-block]}
   \plc{...}
   }
\end{ompcPragma}

where \plc{clause} is one of the following:

{}
\begin{indentedcodelist}
private(\plc{list})
firstprivate(\plc{list})
lastprivate(\plc{[ lastprivate-modifier}:\plc{] list})
reduction(\plc{[reduction-modifier },\plc{] reduction-identifier }:\plc{ list})
allocate(\plc{[allocator }:\plc{] list})
nowait
\end{indentedcodelist}
\end{ccppspecific}

\needspace{16\baselineskip}
\begin{fortranspecific}
The syntax of the \code{sections} construct is as follows:

\begin{ompfPragma}
!$omp sections \plc{[clause[ [},\plc{] clause] ... ]}
   \plc{[}!$omp section\plc{]}
      \plc{structured-block}
   \plc{[}!$omp section
      \plc{structured-block]}
   \plc{...}
!$omp end sections \plc{[}nowait\plc{]}
\end{ompfPragma}

\begin{samepage}
where \plc{clause} is one of the following:

{}
\begin{indentedcodelist}
private(\plc{list})
firstprivate(\plc{list})
lastprivate(\plc{[ lastprivate-modifier}:\plc{] list})
reduction(\plc{[reduction-modifier },\plc{] reduction-identifier }:\plc{ list})
allocate(\plc{[allocator }:\plc{] list})
\end{indentedcodelist}
\end{samepage}
\end{fortranspecific}

\binding
The binding thread set for a \code{sections} region is the current team. A 
\code{sections} region binds to the innermost enclosing \code{parallel} region. 
Only the threads of the team that executes the binding \code{parallel} region 
participate in the execution of the structured blocks and the implied barrier 
of the \code{sections} region if the barrier is not eliminated by a 
\code{nowait} clause.

\descr
Each structured block in the \code{sections} construct is preceded by a 
\code{section} directive except possibly the first block, for which a preceding 
\code{section} directive is optional.

The method of scheduling the structured blocks among the threads in the team is
implementation defined.

There is an implicit barrier at the end of a \code{sections} construct unless a 
\code{nowait} clause is specified.

\events

The \plc{section-begin} event occurs after an implicit task encounters a
\code{sections} construct but before the task executes any 
structured block of the \code{sections} region.

The \plc{sections-end} event occurs after an implicit task finishes execution 
of a \code{sections} region but before it resumes execution of the enclosing context.

The \plc{section-begin} event occurs before an implicit task starts
to execute a structured block in the \code{sections} construct for
each structured block that the task executes.

\omptWorksharingTools{sections}{ompt_work_sections}

A thread dispatches a registered \code{ompt_callback_dispatch}
callback for each occurrence of a \plc{section-begin} 
event in that thread. The callback occurs in the
context of the implicit task.  The callback has type signature
\code{ompt_callback_dispatch_t}. 

\restrictions
Restrictions to the \code{sections} construct are as follows:

\begin{itemize}
\item Orphaned \code{section} directives are prohibited. That is, the 
      \code{section} directives must appear within the \code{sections} 
      construct and must not be encountered elsewhere in the \code{sections} region.
\item The code enclosed in a \code{sections} construct must be a structured block.
\item Only a single \code{nowait} clause can appear on a \code{sections} directive.

\begin{cppspecific}
\item A throw executed inside a \code{sections} region must cause execution to 
      resume within the same section of the \code{sections} region, and the same 
      thread that threw the exception must catch it.
\end{cppspecific}
\end{itemize}

\crossreferences
\begin{itemize}
\item \code{allocate} clause, see
\specref{subsec:allocate Clause}.

\item \code{private}, \code{firstprivate}, \code{lastprivate}, and 
\code{reduction} clauses, see \specref{subsec:Data-Sharing Attribute Clauses}.

\item \code{ompt_scope_begin} and \code{ompt_scope_end}, see
  \specref{sec:ompt_scope_endpoint_t}.

\item \code{ompt_work_sections}, see \specref{sec:ompt_work_t}.

\item \code{ompt_callback_work_t}, see
\specref{sec:ompt_callback_work_t}.

\item \code{ompt_callback_dispatch_t}, see 
\specref{sec:ompt_callback_dispatch_t}.

\end{itemize}



\subsection{\hcode{single} Construct}
\index{single@{\code{single}}}
\index{constructs!single@{\code{single}}}
\label{subsec:single Construct}
\summary
The \code{single} construct specifies that the associated structured block is 
executed by only one of the threads in the team (not necessarily the master 
thread), in the context of its implicit task. The other threads in the team, 
which do not execute the block, wait at an implicit barrier at the end of the 
\code{single} construct unless a \code{nowait} clause is specified.

\syntax
\begin{ccppspecific}
The syntax of the single construct is as follows:

\begin{ompcPragma}
#pragma omp single \plc{[clause[ [},\plc{] clause] ... ] new-line}
   \plc{structured-block}
\end{ompcPragma}

\begin{samepage}
where \plc{clause} is one of the following:

\begin{indentedcodelist}
private(\plc{list})
firstprivate(\plc{list})
copyprivate(\plc{list})
allocate(\plc{[allocator }:\plc{] list})
nowait
\end{indentedcodelist}
\end{samepage}
\end{ccppspecific}

\begin{fortranspecific}
The syntax of the \code{single} construct is as follows:

\begin{ompfPragma}
!$omp single \plc{[clause[ [},\plc{] clause] ... ]}
   \plc{structured-block}
!$omp end single \plc{[end_clause[ [},\plc{] end_clause] ... ]}
\end{ompfPragma}

where \plc{clause} is one of the following:

\begin{indentedcodelist}
private(\plc{list})
firstprivate(\plc{list})
allocate(\plc{[allocator }:\plc{] list})
\end{indentedcodelist}

and \plc{end_clause} is one of the following:

\begin{indentedcodelist}
copyprivate(\plc{list})
nowait
\end{indentedcodelist}
\end{fortranspecific}

\binding
The binding thread set for a \code{single} region is the current team. 
A \code{single} region binds to the innermost enclosing \code{parallel} 
region. Only the threads of the team that executes the binding \code{parallel} 
region participate in the execution of the structured block and the implied 
barrier of the \code{single} region if the barrier is not eliminated by a
\code{nowait} clause.

\descr
Only one of the encountering threads will execute the structured block 
associated with the \code{single} construct. The method of choosing a thread 
to execute the structured block each time the team encounters the construct
is implementation defined. There is an implicit barrier at the end of the 
\code{single} construct unless a \code{nowait} clause is specified.

\events

The \plc{single-begin} event occurs after an \code{implicit task} encounters a
\code{single} construct but before the task starts to execute the structured
block of the \code{single} region.

The \plc{single-end} event occurs after an implicit task finishes execution of 
a \code{single} region but before it resumes execution of the enclosing region.

\tools

A thread dispatches a registered \code{ompt_callback_work} callback with 
\code{ompt_scope_begin} as its \plc{endpoint} argument for each 
occurrence of a \plc{single-begin} event  in that thread. Similarly, a 
thread dispatches a registered \code{ompt_callback_work} callback  with 
\code{ompt_scope_begin} as its \plc{endpoint} argument for each occurrence 
of a \plc{single-end} event in that thread. For each of these callbacks,
the \plc{wstype} argument is \code{ompt_work_single_executor} if the thread
executes the structured block associated with the \code{single} region;
otherwise, the \plc{wstype} argument is  \code{ompt_work_single_other}.
The callback has type signature \code{ompt_callback_work_t}.

\restrictions
Restrictions to the \code{single} construct are as follows:

\begin{itemize}
\item The \code{copyprivate} clause must not be used with the \code{nowait} clause.

\item At most one \code{nowait} clause can appear on a \code{single} construct.

\begin{cppspecific}
\item A throw executed inside a \code{single} region must cause execution to resume within the
same \code{single} region, and the same thread that threw the exception must catch it.
\end{cppspecific}
\end{itemize}


\crossreferences
\begin{itemize}
\item \code{allocate} clause, see
\specref{subsec:allocate Clause}.

\item \code{private} and \code{firstprivate} clauses, see
\specref{subsec:Data-Sharing Attribute Clauses}.

\item \code{copyprivate} clause, see
\specref{subsubsec:copyprivate clause}.

\item \code{ompt_scope_begin} and \code{ompt_scope_end}, see
  \specref{sec:ompt_scope_endpoint_t}.

\item \code{ompt_work_single_executor} and \code{ompt_work_single_other}, see
\specref{sec:ompt_work_t}.

\item \code{ompt_callback_work_t},
\specref{sec:ompt_callback_work_t}.

\end{itemize}














% Here we need to force the blue marker lower, and force the subsection header higher
% in order to reduce the space between the marker and the header, per Richard:
%\begin{samepage}

\begin{fortranspecific}

\subsection{\hcode{workshare} Construct}
\index{workshare@{\code{workshare}}}
\index{constructs!workshare@{\code{workshare}}}
\label{subsec:workshare Construct}
\summary
The \code{workshare} construct divides the execution of the enclosed structured block into
separate units of work, and causes the threads of the team to share the work such that
each unit is executed only once by one thread, in the context of its implicit task.
%\end{samepage}

%\begin{samepage}
\syntax
The syntax of the \code{workshare} construct is as follows:

\begin{ompfPragma}
!$omp workshare
    \plc{structured-block}
!$omp end workshare \plc{[}nowait\plc{]}
\end{ompfPragma}
%\end{samepage}

The enclosed structured block must consist of only the following:

\begin{itemize}
\item array assignments

\item scalar assignments

\item \code{FORALL} statements

\item \code{FORALL} constructs

\item \code{WHERE} statements

\item \code{WHERE} constructs

\item \code{atomic} constructs

\item \code{critical} constructs

\item \code{parallel} constructs
\end{itemize}

Statements contained in any enclosed \code{critical} construct are also subject to these
restrictions. Statements in any enclosed \code{parallel} construct are not restricted.

\binding
The binding thread set for a \code{workshare} region is the current team. A \code{workshare}
region binds to the innermost enclosing \code{parallel} region. Only the threads of the team
executing the binding \code{parallel} region participate in the execution of the units of
work and the implied barrier of the \code{workshare} region if the barrier is not eliminated
by a \code{nowait} clause.

\descr
There is an implicit barrier at the end of a \code{workshare} construct unless a \code{nowait}
clause is specified.

An implementation of the \code{workshare} construct must insert any synchronization that is
required to maintain standard Fortran semantics. For example, the effects of one
statement within the structured block must appear to occur before the execution of
succeeding statements, and the evaluation of the right hand side of an assignment must
appear to complete prior to the effects of assigning to the left hand side.

The statements in the \code{workshare} construct are divided into units of work as follows:

\begin{itemize}
\item For array expressions within each statement, including transformational array
intrinsic functions that compute scalar values from arrays:

\begin{itemize} % nested level
\item Evaluation of each element of the array expression, including any references to
\code{ELEMENTAL} functions, is a unit of work.

\item Evaluation of transformational array intrinsic functions may be freely subdivided
into any number of units of work.
\end{itemize}

\item For an array assignment statement, the assignment of each element is a unit of work.

\item For a scalar assignment statement, the assignment operation is a unit of work.

\item For a \code{WHERE} statement or construct, the evaluation of the mask expression and the
masked assignments are each a unit of work.

\item For a \code{FORALL} statement or construct, the evaluation of the mask expression,
expressions occurring in the specification of the iteration space, and the masked
assignments are each a unit of work.

\item For an \code{atomic} construct, the atomic operation on the storage location designated as
\plc{x} is a unit of work.

\item For a \code{critical} construct, the construct is a single unit of work.

\item For a \code{parallel} construct, the construct is a unit of work with respect to the
\code{workshare} construct. The statements contained in the \code{parallel} construct are
executed by a new thread team.

\item If none of the rules above apply to a portion of a statement in the structured block,
then that portion is a unit of work.
\end{itemize}

The transformational array intrinsic functions are \code{MATMUL}, \code{DOT_PRODUCT}, \code{SUM},
\code{PRODUCT}, \code{MAXVAL}, \code{MINVAL}, \code{COUNT},
\code{ANY}, \code{ALL}, \code{SPREAD}, \code{PACK}, \code{UNPACK},
\code{RESHAPE}, \code{TRANSPOSE}, \code{EOSHIFT}, \code{CSHIFT}, \code{MINLOC}, and \code{MAXLOC}.

It is unspecified how the units of work are assigned to the threads executing a
\code{workshare} region.

If an array expression in the block references the value, association status, or allocation
status of private variables, the value of the expression is undefined, unless the same
value would be computed by every thread.

If an array assignment, a scalar assignment, a masked array assignment, or a \code{FORALL}
assignment assigns to a private variable in the block, the result is unspecified.

The \code{workshare} directive causes the sharing of work to occur only in the \code{workshare}
construct, and not in the remainder of the \code{workshare} region.

%\tools
\omptWorksharing{workshare}{ompt_work_workshare}

\begin{samepage}
\restrictions
The following restrictions apply to the \code{workshare} construct:

\begin{itemize}
\item All array assignments, scalar assignments, and masked array assignments must be
intrinsic assignments.

\item The construct must not contain any user defined function calls unless the function is
\code{ELEMENTAL}.
\end{itemize}
\end{samepage}

\crossreferences
\begin{itemize}

\item \code{atomic} construct, see \specref{subsec:atomic Construct}.

\item \code{critical} construct, see \specref{subsec:critical Construct}.

\item \code{parallel} construct, see \specref{sec:parallel Construct}.

\item \code{ompt_scope_begin} and \code{ompt_scope_end}, see
  \specref{sec:ompt_scope_endpoint_t}.
\item \code{ompt_work_workshare}, see \specref{sec:ompt_work_t}.
\item \code{ompt_callback_work_t}, see
\specref{sec:ompt_callback_work_t}.
\end{itemize}

%\filbreak
\end{fortranspecific}
